### 2013-12-06 paper.tex
added the paper latex file

\documentclass[10pt]{article}
\usepackage{fullpage}
\usepackage[pdftex]{graphicx}
\usepackage{rotating}
\usepackage{multirow}
\usepackage{subfigure}
\usepackage{url}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{wasysym}

\newcommand{\etal}{\emph{et al.}}

\begin{document}
\title{K Nearest Neighbour Search Using Space-filling Curves}
\author{Camara Lerner \hspace{2cm} Christopher Rabl \\
  Department of Mathematics and Computer Science\\
  University of Lethbridge, Canada}

\maketitle

\begin{abstract}

\end{abstract}

\section{Introduction}

\section{Hilbert Curve}

Formula for Gray Code:
$ gc(i) = i \oplus \left( i \gg 1 \right)$

Formula for $g(i) = \lfloor \log _{2} \left( gc(i) \oplus gc(i+1) \right) \rfloor + 1$

Formula for
\[ d(i) = \left\{ 
  \begin{array}{l l}
    0 & \quad \text{if $i = 0$}\\
    g(i - 1) \text{ mod } n & \quad \text{if $i$ is even} \\
    g(i) \text{ mod }  n & \quad \text{if $i$ is odd}
  \end{array} \right.\]

Formula for the bitwise right rotation:
$a \rightturn i = \left[ a_{n - 1 + i\text{ mod }n} \ldots a_{i\text{ mod }n} \right]_{\left[ 2 \right]}, \text{ where } a = \left[ a_{n-1} \ldots a_0\right]_{\left[ 2 \right]} $

The bitwise left rotation is similar.

Formula for 
\[ e(i) = \left\{ 
  \begin{array}{l l}
    0 & \quad \text{if $i = 0$}\\
    e(i - 1) \oplus 2^{d(i-1)} \oplus 2^{g(i-1)}& \quad \text{otherwise} \\
  \end{array} \right.\]

\begin{table}

  \begin{tabular}{l}
    \hline
    Algorithm for Gray Code Inverse $\left( gc^{-1}\left( g \right) \right)$\\
    \hline
    Given $g \in \mathbb{N}$, calculates the $i \in \mathbb{N}$ such that $gc\left( i \right) = g $ \\
    {\bf Input:} $g \in \mathbb{N}$\\
    {\bf Output:} $i \in \mathbb{N}$ such that $gc \left( i \right) = g $\\
    1: $ m \leftarrow $ number of bits to represent $g$ \\
    2: $ \left( i, j \right) \leftarrow \left( g, 1 \right) $ \\
    3: {\bf while} $ j < m $ {\bf do} \\
    4: {     } $ i \leftarrow i \oplus \left( g \gg j \right)$ \\
    5: {     } $ j \leftarrow j + 1$ \\
    6: {\bf end while} \\
    \hline
    % \oplus is exclusive or
    % \ll left shift
    % \gg right shift
    % \mid bitwise or
  \end{tabular}
  \label{gray-code-inverse}
\end{table}

\begin{table}

  \begin{tabular}{l}
    \hline
    Algorithm for Hilbert Index $(hi(n, m, {\bf p}))$ \\
    \hline
    Calculates the Hilbert index corresponding to the point ${\bf p} \in \mathbb{N}^n $ \\
    {\bf Input:} $n, m \in \mathbb{N} - \{0\} $ and a point ${\bf p} \in \mathbb{N}^n$ \\
    {\bf Output:} $h \in \mathbb{N}$, the Hilbert index of the point ${\bf p}$. \\
    1: $ \left( h, e, d \right) \leftarrow \left( 0, 0, 0 \right) $ \\
    2: {\bf for} $ i = m - 1 $ to $ 0 $ { \bf do} \\
    3: {     } $ l \leftarrow \left[ \left( p_{n-1} , i \right) \ldots \left( p_0 , i \right) \right]_{\left[ 2 \right]} $ \\
    4: {     } $ l \leftarrow \left( b \oplus e \right) \rightturn \left( d+1 \right)$ \\
    5: {     } $ w \leftarrow gc^{-1} \left( l \right)$ \\
    6: {     } $ e \leftarrow e \oplus \left( e \left( w \right) \leftturn \left( d+ 1 \right) \right) $ \\
    7: {     } $ d \leftarrow d + d \left( w \right) + 1 $ mod $n$ \\
    8: {     } $ h \leftarrow \left( h \ll n \right) | w $ \\
    9: {\bf end for} \\
    \hline
  \end{tabular}
  \label{hilbert-point-to-index}
\end{table}

\end{document}

### 2013-12-06 src/main.py
Added average correctness stuff

@@ -153,8 +153,10 @@ def main():
     P = [(x,y,z) for x in range(10) for y in range(10) for z in range(10)]
     t_cum_aknn = 0
     t_cum_knn = 0
-    k = 6
+    k = 26
     precision = 32
+
+    average_correct = 0
     for i in range(len(P)):
         t0_aknn = time.time()
         AKNN_i = construct_morton(P, i, k)
@@ -165,11 +167,16 @@ def main():
         KNN_i = set_knn(P[i], k, P)
         tf_knn = time.time() - t0_knn
         t_cum_knn += tf_knn
-        print P[i], "\tAkNN:", AKNN_i, "\tOPT:", KNN_i
+        
+        num_correct = sum([item in AKNN_i for item in KNN_i])
+        average_correct += num_correct
+        print P[i], "\tCorrect:", num_correct, "/", len(KNN_i)
     print ""
-    print "Time AKNN:", t_cum_aknn
+    print "Time Morton AKNN:", t_cum_aknn
     print "Time OPT:", t_cum_knn
+    print "Average Correct:", average_correct / len(P)
 
+    average_correct = 0
     for i in range(len(P)):
         t0_aknn = time.time()
         AKNN_i = construct_hilbert(P, i, k, precision)
@@ -180,10 +187,14 @@ def main():
         KNN_i = set_knn(P[i], k, P)
         tf_knn = time.time() - t0_knn
         t_cum_knn += tf_knn
-        print P[i], "\tAkNN:", AKNN_i, "\tOPT:", KNN_i
+        
+        num_correct = sum([item in AKNN_i for item in KNN_i])
+        average_correct += num_correct
+        print P[i], "\tCorrect:", num_correct, "/", len(KNN_i)
     print ""
-    print "Time AKNN:", t_cum_aknn
+    print "Time Hilbert AKNN:", t_cum_aknn
     print "Time OPT:", t_cum_knn
+    print "Average Correct:", average_correct / len(P)
 
 
 if __name__ == "__main__":

### 2013-12-06 src/morton.py

@@ -1,6 +1,8 @@
 #!/usr/bin/env python
 # Stolen from Wikipedia
 
+import numpy as np
+
 def less_msb(x, y):
     return x < y and x < (x ^ y)
 

### 2013-12-06 paper.tex
added bib file. Cannot get it to compile just yet

@@
 
-Formula for Gray Code:
-$ gc(i) = i \oplus \left( i \gg 1 \right)$
+The Hilbert Curve is a space filling curve that makes the next point in the curve only one unit distance from the previous point.
+The algorithm for calculating the Hilbert Curve used is found in~\cite{Hamilton:2006}.
 
-Formula for $g(i) = \lfloor \log _{2} \left( gc(i) \oplus gc(i+1) \right) \rfloor + 1$
+\flushleft
+Formula for Gray Code is in \ref{gray-code}.
+\begin{equation}
+  \label{gray-code}
+  gc(i) = i \oplus \left( i \gg 1 \right) 
+\end{equation}
 
-Formula for
-\[ d(i) = \left\{ 
-  \begin{array}{l l}
+Formula for $g$ is displayed in \ref{g}.
+\begin{equation}
+  \label{g}
+  g(i) = \lfloor \log _{2} \left( gc(i) \oplus gc(i+1) \right) \rfloor + 1
+\end{equation}
+
+Formula for $d(i)$ is displayed in \ref{d}.
+\begin{equation}
+  \label{d}
+  d(i) = 
+    \begin{cases}
       0 & \quad \text{if $i = 0$}\\
@@
       g(i) \text{ mod }  n & \quad \text{if $i$ is odd}
-  \end{array} \right.\]
+    \end{cases}
+\end{equation}
 
-Formula for the bitwise right rotation:
-$a \rightturn i = \left[ a_{n - 1 + i\text{ mod }n} \ldots a_{i\text{ mod }n} \right]_{\left[ 2 \right]}, \text{ where } a = \left[ a_{n-1} \ldots a_0\right]_{\left[ 2 \right]} $
+Formula for the bitwise right rotation is displayed in \ref{right-rotate}. The bitwise left rotation is similar.
+\begin{equation}
+  \label{right-rotate}
+  a \rightturn i = \left[ a_{n - 1 + i\text{ mod }n} \ldots a_{i\text{ mod }n} \right]_{\left[ 2 \right]}, \text{ where } a = \left[ a_{n-1} \ldots a_0\right]_{\left[ 2 \right]} 
+\end{equation}
 
-The bitwise left rotation is similar.
-
-Formula for 
-\[ e(i) = \left\{ 
-  \begin{array}{l l}
+Formula for $e$ is displayed in \ref{e}.
+\begin{equation}
+  \label{e}
+  e(i) =  
+    \begin{cases}
       0 & \quad \text{if $i = 0$}\\
       e(i - 1) \oplus 2^{d(i-1)} \oplus 2^{g(i-1)}& \quad \text{otherwise} \\
-  \end{array} \right.\]
+    \end{cases}
+\end{equation}
 
 \begin{table}
-
+  \flushleft
   \begin{tabular}{l}
-    \hline
-    Algorithm for Gray Code Inverse $\left( gc^{-1}\left( g \right) \right)$\\
     \hline
@@
     3: {\bf while} $ j < m $ {\bf do} \\
-    4: {     } $ i \leftarrow i \oplus \left( g \gg j \right)$ \\
-    5: {     } $ j \leftarrow j + 1$ \\
+    4: \qquad $ i \leftarrow i \oplus \left( g \gg j \right)$ \\
+    5: \qquad $ j \leftarrow j + 1$ \\
     6: {\bf end while} \\
@@
   \end{tabular}
+  \caption{The algorithm for calculating the Gray Code Inverse. }
   \label{gray-code-inverse}
@@
     \hline
-    Algorithm for Hilbert Index $(hi(n, m, {\bf p}))$ \\
-    \hline
     Calculates the Hilbert index corresponding to the point ${\bf p} \in \mathbb{N}^n $ \\
@@
     2: {\bf for} $ i = m - 1 $ to $ 0 $ { \bf do} \\
-    3: {     } $ l \leftarrow \left[ \left( p_{n-1} , i \right) \ldots \left( p_0 , i \right) \right]_{\left[ 2 \right]} $ \\
-    4: {     } $ l \leftarrow \left( b \oplus e \right) \rightturn \left( d+1 \right)$ \\
-    5: {     } $ w \leftarrow gc^{-1} \left( l \right)$ \\
-    6: {     } $ e \leftarrow e \oplus \left( e \left( w \right) \leftturn \left( d+ 1 \right) \right) $ \\
-    7: {     } $ d \leftarrow d + d \left( w \right) + 1 $ mod $n$ \\
-    8: {     } $ h \leftarrow \left( h \ll n \right) | w $ \\
+    3: \qquad $ l \leftarrow \left[ \left( p_{n-1} , i \right) \ldots \left( p_0 , i \right) \right]_{\left[ 2 \right]} $ \\
+    4: \qquad $ l \leftarrow \left( b \oplus e \right) \rightturn \left( d+1 \right)$ \\
+    5: \qquad $ w \leftarrow gc^{-1} \left( l \right)$ \\
+    6: \qquad $ e \leftarrow e \oplus \left( e \left( w \right) \leftturn \left( d+ 1 \right) \right) $ \\
+    7: \qquad $ d \leftarrow d + d \left( w \right) + 1 $ mod $n$ \\
+    8: \qquad $ h \leftarrow \left( h \ll n \right) | w $ \\
     9: {\bf end for} \\
@@
   \end{tabular}
+  \caption{The algorithm for calculating the hilbert index given any dimensional point as long as the bits used within an index of the hilbert curve space is specified in ($m$). }
   \label{hilbert-point-to-index}

### 2013-12-06 src/main.py
Added new measure of badness

@@ -10,11 +10,17 @@ import numpy
 import sys
 import math
 import bitstring
-import numpy as np
+import numpy as np 
 import time
+import warnings
+
+# Ignore exponent divide-by-zero warning
+warnings.simplefilter("ignore", RuntimeWarning) 
 
 from hilbert import *
 from morton import *
+
+
 
 V_CONSTANT = 4 # Platform-dependent constant
 
@@ -149,15 +155,26 @@ def construct_hilbert(points, i, k, precision):
     return A_i
         
 
-def main():
-    P = [(x,y,z) for x in range(10) for y in range(10) for z in range(10)]
+def main(file_name, k):
+    #P = [(x,y) for x in range(20) for y in range(20)]
+    dataset = np.genfromtxt(str(file_name), delimiter="\t")
+    P = [tuple(item) for item in dataset]
+
     t_cum_aknn = 0
     t_cum_knn = 0
-    k = 26
+    k = int(k)
     precision = 32
 
-    average_correct = 0
+    print "Running Z-order AKNN comparison on", len(P), "points..."
+    average_distance_from_opt = 0
     for i in range(len(P)):
+        iter_fraction = float(i) / len(P)
+
+        # Progress bar
+        amtDone = iter_fraction * 100
+        sys.stdout.write("\r%.1f%%" %amtDone)
+        sys.stdout.flush()
+
         t0_aknn = time.time()
         AKNN_i = construct_morton(P, i, k)
         tf_aknn = time.time() - t0_aknn
@@ -168,16 +185,25 @@ def main():
         tf_knn = time.time() - t0_knn
         t_cum_knn += tf_knn
         
-        num_correct = sum([item in AKNN_i for item in KNN_i])
-        average_correct += num_correct
-        print P[i], "\tCorrect:", num_correct, "/", len(KNN_i)
+        average_distance_from_opt += set_radius(AKNN_i) / set_radius(KNN_i)
     print ""
     print "Time Morton AKNN:", t_cum_aknn
     print "Time OPT:", t_cum_knn
-    print "Average Correct:", average_correct / len(P)
+    print "Average times distance from OPT:", average_distance_from_opt / len(P)
 
-    average_correct = 0
+    t_cum_aknn = 0
+    t_cum_knn = 0
+    print ""
+    print "Running Hilbert AKNN comparison on", len(P), "points..."
+    average_distance_from_opt = 0
     for i in range(len(P)):
[…]

### 2013-12-06 paper.tex
added the algorithms correctly

@@
 \usepackage{wasysym}
+\usepackage{algorithm}
+\usepackage{algpseudocode}
 
@@
 The Hilbert Curve is a space filling curve that makes the next point in the curve only one unit distance from the previous point.
-The algorithm for calculating the Hilbert Curve used is found in~\cite{Hamilton:2006}.
+The curve is defined based onthe amount of space the curve fills.
+The space that the curve fills can be in any dimensionally space $(n)$ with the length of each dimension filled equal to $2^m$.
+This means that for each point in the curve the maximum bit size for each dimension is $(m)$~\cite{Hamilton:2006}.
+
+The curve labels the points from $0$ to $2^{m*n} - 1$ in the order that the curve visits each point.
+The Hilbert Curve considered here will start in the bottom left hand corner and finishes in the upper left corner.
 
@@
 
-\begin{table}
-  \flushleft
-  \begin{tabular}{l}
-    \hline
-    Given $g \in \mathbb{N}$, calculates the $i \in \mathbb{N}$ such that $gc\left( i \right) = g $ \\
-    {\bf Input:} $g \in \mathbb{N}$\\
-    {\bf Output:} $i \in \mathbb{N}$ such that $gc \left( i \right) = g $\\
-    1: $ m \leftarrow $ number of bits to represent $g$ \\
-    2: $ \left( i, j \right) \leftarrow \left( g, 1 \right) $ \\
-    3: {\bf while} $ j < m $ {\bf do} \\
-    4: \qquad $ i \leftarrow i \oplus \left( g \gg j \right)$ \\
-    5: \qquad $ j \leftarrow j + 1$ \\
-    6: {\bf end while} \\
-    \hline
+\begin{algorithm}
+  \caption{The algorithm for calculating the Gray Code Inverse.     Given $g \in \mathbb{N}$, calculates the $i \in \mathbb{N}$ such that $gc\left( i \right) = g $ }
+  \label{gray-code-inverse}
+  \begin{algorithmic}[1]
+    \Require $g \in \mathbb{N}$
+    \Return $i \in \mathbb{N} \text{ such that } gc \left( i \right) = g $
+    \State $ m \leftarrow \text{ number of bits to represent } g$ 
+    \State $ \left( i, j \right) \leftarrow \left( g, 1 \right) $ 
+    \While{$ j < m $ }
+    \State $ i \leftarrow i \oplus \left( g \gg j \right)$ 
+    \State $ j \leftarrow j + 1$ 
+    \EndWhile
     % \oplus is exclusive or
@@
     % \mid bitwise or
-  \end{tabular}
-  \caption{The algorithm for calculating the Gray Code Inverse. }
-  \label{gray-code-inverse}
-\end{table}
+  \end{algorithmic}
+\end{algorithm}
 
-\begin{table}
+\begin{algorithm}
+  \caption{Calculating the Hilbert index given any dimensional point ${\bf p}$ 
+    of size $n$ as long as the bits used within an index of the hilbert curve 
+    space is specified in ($m$).}
+  \label{hilbert-point-to-index}
+  \begin{algorithmic}[1]
+    \Require $n, m \in \mathbb{N} - \{0\} \text{ and a point } {\bf p} \in \mathbb{N}^n$
+    \Return $h \in \mathbb{N} \text{, the Hilbert index of the point }{\bf p}$
+    \State $ \left( h, e, d \right) \leftarrow \left( 0, 0, 0 \right) $
+    \For{$j\gets m - 1, 0$} 
+    \State $ l \leftarrow \left[ \left( p_{n-1} , i \right) \ldots \left( p_0 , i \right) \right]_{\left[ 2 \right]} $ 
+    \State $ l \leftarrow \left( b \oplus e \right) \rightturn \left( d+1 \right)$ 
+    \State $ w \leftarrow gc^{-1} \left( l \right)$ 
+    \State $ e \leftarrow e \oplus \left( e \left( w \right) \leftturn \left( d+ 1 \right) \right) $ 
+    \State $ d \leftarrow d + d \left( w \right) + 1 \text{ mod }n$ 
+    \State $ h \leftarrow \left( h \ll n \right) | w $
+    \EndFor
+  \end{algorithmic}
+\end{algorithm}
 
-  \begin{tabular}{l}
-    \hline
-    Calculates the Hilbert index corresponding to the point ${\bf p} \in \mathbb{N}^n $ \\
-    {\bf Input:} $n, m \in \mathbb{N} - \{0\} $ and a point ${\bf p} \in \mathbb{N}^n$ \\
-    {\bf Output:} $h \in \mathbb{N}$, the Hilbert index of the point ${\bf p}$. \\
-    1: $ \left( h, e, d \right) \leftarrow \left( 0, 0, 0 \right) $ \\
-    2: {\bf for} $ i = m - 1 $ to $ 0 $ { \bf do} \\
-    3: \qquad $ l \leftarrow \left[ \left( p_{n-1} , i \right) \ldots \left( p_0 , i \right) \right]_{\left[ 2 \right]} $ \\
-    4: \qquad $ l \leftarrow \left( b \oplus e \right) \rightturn \left( d+1 \right)$ \\
-    5: \qquad $ w \leftarrow gc^{-1} \left( l \right)$ \\
-    6: \qquad $ e \leftarrow e \oplus \left( e \left( w \right) \leftturn \left( d+ 1 \right) \right) $ \\
-    7: \qquad $ d \leftarrow d + d \left( w \right) + 1 $ mod $n$ \\
-    8: \qquad $ h \leftarrow \left( h \ll n \right) | w $ \\
-    9: {\bf end for} \\
-    \hline
-  \end{tabular}
-  \caption{The algorithm for calculating the hilbert index given any dimensional point as long as the bits used within an index of the hilbert curve space is specified in ($m$). }
-  \label{hilbert-point-to-index}
-\end{table}
+\bibliographystyle{amsplain}
+\bibliography{mybiblio}
 

### 2013-12-06 src/main.py
Forgot to add hilbert

@@ -164,7 +164,7 @@ def main(file_name, k):
     t_cum_knn = 0
     k = int(k)
     precision = 32
-
+    """
     print "Running Z-order AKNN comparison on", len(P), "points..."
     average_distance_from_opt = 0
     for i in range(len(P)):
@@ -190,7 +190,9 @@ def main(file_name, k):
     print "Time Morton AKNN:", t_cum_aknn
     print "Time OPT:", t_cum_knn
     print "Average times distance from OPT:", average_distance_from_opt / len(P)
-
+    """
+    print "Constructing Hilbert Ordering..."
+    hilbert_index(P, len(P[0]), precision)
     t_cum_aknn = 0
     t_cum_knn = 0
     print ""

### 2013-12-06 src/hilbert.py
Added unit tests and new point class

@@ -1,110 +1,35 @@
-#!/usr/bin/env python
-# Hilbert Keying Algo
-# Camara Lerner - University of Lethbridge
+import utils, math
 
-#run source venv
-import math
-import bitstring
-import numpy as np
+def transform(entry, direction, width, x):
+    assert x < 2**width
+    assert entry < 2**width
+    return utils.rrot((x^entry), direction+1, width)
 
-HILBERT_ORDERED = []
-
-# gray code
-def gc(i): 
-    return i ^ (i >> 1)
-
-# gray code Inverse
-def gcInverse(g):
-    if g == 0:
+def direction(x, n):
+    assert x < 2**n
+    if x == 0:
         return 0
-    m = int(math.floor(math.log(g, 2))+1) # number of bits required to represent g
-    i = g
-    j = 1
-    while j < m:
-        i = i ^ (g >> j)
-        j =  j + 1
-    return i
-
-# does a left bitwise rotation of a by b bits, with the number of bits in a 
-#    being m
-def leftRotate(a, b, m):
-    w = bitstring.BitArray(uint = a, length = m)
-    w.rol(b)
-    return int(w.uint)
-
-def rightRotate(a, b, m):
-    w = bitstring.BitArray(uint = a, length = m)
-    w.ror(b)
-    return int(w.uint)
-
-def g(i):
-    return int(math.log(gc(i) ^ gc(i+1), 2))
-
-def D(i, n):
-    if i < 0:
-        print "d: WHY IS THERE A NEGATIVE NUMBER >:|"
-        return 0
-    elif i == 0:
-        return 0
-    elif (i % 2) == 0:
-        return g(i - 1) % n
-    elif (i % 2) == 1:
-        return g(i) % n
+    elif x%2 == 0:
+        return utils.tsb(x-1, n)%n
     else:
-        print "d: FIX ME!!!!!!"
+        return utils.tsb(x, n)%n
+
+def entry(x):
+    if x == 0:
         return 0
+    else:
+        return utils.graycode(2*((x-1)/2))
 
-def E(i, n):
-    if i == 0:
-        return 0
-    return E(i - 1, n) ^ (2 ** D(i - 1, n)) ^ (2 ** g(i - 1))
-
-def L(i, p, n, m):
-    l = bitstring.BitArray()
-    for j in range(0, n):
-        q = bitstring.BitArray(uint = p[j], length = m)
-        q.reverse()
-        l += q[i:i + 1]
[…]

### 2013-12-06 src/point.py

@@ -1,0 +1,33 @@
+import math
+
+import hilbert
+import morton
+
+class Point(object):
+    def __init__(self, coordinates, precision=32, sftype="hilbert"):
+        self.coordinates = coordinates # Tuple of point coordinates
+        self.dimension = len(coordinates)
+        self.precision = precision
+        self.sftype = sftype.lower()
+        if not self.sftype in ["hilbert", "morton"]:
+            raise ValueError("Point type must be either Hilbert or Morton")
+        
+    def index_hilbert(self):
+        return hilbert.hilbert_index(self.dimension, self.precision, self.coordinates)
+    
+    def __lt__(self, other):
+        if self.sftype == "hilbert":
+            return self.index_hilbert() < other.index_hilbert()
+        return morton.cmp_zorder(self.coordinates, other.coordinates) < 0
+
+    def __eq__(self, other):
+        if self.sftype == "hilbert":
+            return self.index_hilbert() == other.index_hilbert()
+        return morton.cmp_zorder(self.coordinates, other.coordinates) == 0
+
+    def __pow__(self, power):
+        result = Point(tuple([x ** power for x in self.coordinates]), precision=self.precision, sftype=self.sftype)
+        return result
+
+    def __repr__(self):
+        return str(self.coordinates)

### 2013-12-06 src/tester.py

@@ -1,0 +1,18 @@
+from point import *
+import unittest
+
+class HilbertPointTests(unittest.TestCase):
+    a = Point((1,0,1), sftype='hilbert')
+    b = Point((0,1,1), sftype='hilbert')
+
+    def test_index_comp(self):
+        self.assertEqual(self.a > self.b, False)
+        self.assertEqual(self.a == self.b, False)
+        self.assertEqual(self.a < self.b, True)
+
+    def test_correct_order(self):
+        self.assertEqual(self.a.index_hilbert(), 2)
+        self.assertEqual(self.b.index_hilbert(), 4)
+    
+if __name__ == "__main__":
+    unittest.main()

### 2013-12-06 src/tester.py
Added some more unit tests

@@ -1,6 +1,13 @@
 from point import *
 import unittest
 
+class PointTests(unittest.TestCase):
+    def create_invalid(self):
+        return Point((0,0,0), sftype='invalid')
+    
+    def test_create_invalid(self):
+        self.assertRaises(ValueError, self.create_invalid)
+
 class HilbertPointTests(unittest.TestCase):
     a = Point((1,0,1), sftype='hilbert')
     b = Point((0,1,1), sftype='hilbert')
@@ -13,6 +20,15 @@ class HilbertPointTests(unittest.TestCase):
     def test_correct_order(self):
         self.assertEqual(self.a.index_hilbert(), 2)
         self.assertEqual(self.b.index_hilbert(), 4)
+
+class MortonPointTests(unittest.TestCase):
+    a = Point((1,0,1), sftype='morton')
+    b = Point((1,1,0), sftype='morton')
+    
+    def test_index_comp(self):
+        self.assertEqual(self.a < self.b, True)
+        self.assertEqual(self.a == self.b, False)
+        self.assertEqual(self.a > self.b, False)
     
 if __name__ == "__main__":
     unittest.main()

### 2013-12-06 src/main.py
Point now checks if it is valid

@@ -17,10 +17,7 @@ import warnings
 # Ignore exponent divide-by-zero warning
 warnings.simplefilter("ignore", RuntimeWarning) 
 
-from hilbert import *
-from morton import *
-
-
+from point import *
 
 V_CONSTANT = 4 # Platform-dependent constant
 

### 2013-12-06 src/point.py

@@ -12,15 +12,23 @@ class Point(object):
         if not self.sftype in ["hilbert", "morton"]:
             raise ValueError("Point type must be either Hilbert or Morton")
         
+    def check_valid(self):
+        if any([x < 0 for x in self.coordinates]):
+            raise ValueError("Cannot get Hilbert index of negative point")
+
+
     def index_hilbert(self):
+        self.check_valid()
         return hilbert.hilbert_index(self.dimension, self.precision, self.coordinates)
     
     def __lt__(self, other):
+        self.check_valid()
         if self.sftype == "hilbert":
             return self.index_hilbert() < other.index_hilbert()
         return morton.cmp_zorder(self.coordinates, other.coordinates) < 0
 
     def __eq__(self, other):
+        self.check_valid()
         if self.sftype == "hilbert":
             return self.index_hilbert() == other.index_hilbert()
         return morton.cmp_zorder(self.coordinates, other.coordinates) == 0
@@ -28,6 +36,15 @@ class Point(object):
     def __pow__(self, power):
         result = Point(tuple([x ** power for x in self.coordinates]), precision=self.precision, sftype=self.sftype)
         return result
+
+    def __add__(self, other):
+        return Point(tuple([self[i] - other[i] for i in range(self.dimensions)]), precision=self.precision, sftype=self.sftype)
+    
+    def __sub__(self, other):
+        return Point(tuple([self[i] - other[i] for i in range(self.dimensions)]), precision=self.precision, sftype=self.sftype)
+
+    def sum(self):
+        return sum(self.coordinates)
 
     def __repr__(self):
         return str(self.coordinates)

### 2013-12-07 src/main.py
Shifting operation now works

@@ -15,7 +15,7 @@ import time
 import warnings
 
 # Ignore exponent divide-by-zero warning
-warnings.simplefilter("ignore", RuntimeWarning) 
+#warnings.simplefilter("ignore", RuntimeWarning) 
 
 from point import *
 
@@ -23,7 +23,7 @@ V_CONSTANT = 4 # Platform-dependent constant
 
 # Euclidean distance between two points
 def distance(a, b):
-    return np.sqrt(((np.array(a)-np.array(b))**2).sum())
+    return math.sqrt(((a - b) ** 2).sum())
 
 # Find the indices of the k smallest items in a set S, assuming that S contains a zero element
 def k_smallest_indices(k, S):
@@ -33,10 +33,10 @@ def k_smallest_indices(k, S):
 
 # Find the k nearest neighbors to a point in a given set S
 def set_knn(point, k, S):
-    distance_array = [distance(point,s) for s in S]
+    distances = [distance(point,s) for s in S]
 
     # Find the indices of the k smallest items in S
-    smallest_indices = k_smallest_indices(k, distance_array)
+    smallest_indices = k_smallest_indices(k, distances)
 
     # Find the points corresponding to those indices
     nearest_neighbors = [S[i] for i in smallest_indices]
@@ -47,28 +47,29 @@ def box_dist(point, box_center, box_radius):
 
 # The radius of a set is the largest distance between any two points in the set S
 def set_radius(S):
-    distance_matrix = np.zeros((len(S), len(S)))
-    distance_matrix.flat = [distance(p0, p1) for p0 in S for p1 in S]
-    return np.max(distance_matrix)
+    distances = [distance(p0, p1) for p0 in S for p1 in S]
+    return np.max(distances)
 
-def csearch_morton(point, A_i, i, k, low, hi):
+def csearch(point, A_i, i, k, low, hi):
     if (hi-low) < V_CONSTANT:
         return set_knn(point[i], k, list(set(A_i + point[low:hi])))
+
     mid = int((hi + low) / 2)
-    A_i = set_knn(point[i], k, list(set(A_i + point[mid])))
-    if box_dist(point[i], point[low], dist(point[low], point[hi])) > set_radius(A_i):
+    A_i = set_knn(point[i], k, list(set(A_i + [point[mid]])))
[…]

### 2013-12-07 src/point.py

@@ -4,32 +4,41 @@ import hilbert
 import morton
 
 class Point(object):
-    def __init__(self, coordinates, precision=32, sftype="hilbert"):
-        self.coordinates = coordinates # Tuple of point coordinates
+    def __init__(self, coordinates, precision=32, sftype='hilbert'):
+        self.coordinates = tuple([int(item) for item in coordinates]) # Tuple of point coordinates
         self.dimension = len(coordinates)
         self.precision = precision
         self.sftype = sftype.lower()
-        if not self.sftype in ["hilbert", "morton"]:
-            raise ValueError("Point type must be either Hilbert or Morton")
+        if not self.sftype in ['hilbert', 'morton']:
+            raise ValueError('Point type must be either Hilbert or Morton')
         
     def check_valid(self):
         if any([x < 0 for x in self.coordinates]):
-            raise ValueError("Cannot get Hilbert index of negative point")
-
+            #raise ValueError('Cannot get Hilbert index of negative point')
+            return False
+        return True
 
     def index_hilbert(self):
-        self.check_valid()
+        if not self.check_valid():
+            return -1
         return hilbert.hilbert_index(self.dimension, self.precision, self.coordinates)
+
+    def sum(self):
+        return sum(self.coordinates)
+
+    def shift(self, amount):
+        result = Point(tuple([x + amount for x in self.coordinates]), precision=self.precision, sftype=self.sftype)
+        return result
     
     def __lt__(self, other):
         self.check_valid()
-        if self.sftype == "hilbert":
+        if self.sftype == 'hilbert':
             return self.index_hilbert() < other.index_hilbert()
         return morton.cmp_zorder(self.coordinates, other.coordinates) < 0
 
     def __eq__(self, other):
         self.check_valid()
-        if self.sftype == "hilbert":
+        if self.sftype == 'hilbert':
             return self.index_hilbert() == other.index_hilbert()
         return morton.cmp_zorder(self.coordinates, other.coordinates) == 0
 
@@ -38,13 +47,13 @@ class Point(object):
         return result
 
     def __add__(self, other):
-        return Point(tuple([self[i] - other[i] for i in range(self.dimensions)]), precision=self.precision, sftype=self.sftype)
+        return Point(tuple([self[i] - other[i] for i in range(self.dimension)]), precision=self.precision, sftype=self.sftype)
     
     def __sub__(self, other):
[…]

### 2013-12-07 paper/paper.tex
moving the paper things into the paper folder

--- 2013-12-06 paper.tex
@@
 The Hilbert Curve is a space filling curve that makes the next point in the curve only one unit distance from the previous point.
-The curve is defined based onthe amount of space the curve fills.
+The curve is defined based on the amount of space the curve fills.
 The space that the curve fills can be in any dimensionally space $(n)$ with the length of each dimension filled equal to $2^m$.
@@
 The curve labels the points from $0$ to $2^{m*n} - 1$ in the order that the curve visits each point.
-The Hilbert Curve considered here will start in the bottom left hand corner and finishes in the upper left corner.
+The Hilbert Curve considered here will start in the bottom left hand corner and finishes in the upper left corner~\cite{Hamilton:2006}.
+
+With Hilbert curves that which uses only one bit called the unit Hilbert curve, is reflected and rotated and moved around to fill the space of the larger Hilbert Curve, and then these unit Hilbert curves are connected to make the overall Hilbert curve.
+The only constraint on the unit Hilbert curves is that the start and end points must match up with their neighbouring unit Hilbert curves.
+The same idea is done with gray code, but they  can only be reflected and reversed.
+These facts are used to inverse the operation of and get the point from the index in the list for both gray code and Hilbert indexing.
+Hilbert curves also can use information from gray code in determining the changes to the unit Hilbert curve~\cite{Hamilton:2006}.
 
@@
       0 & \quad \text{if $i = 0$}\\
-      g(i - 1) \text{ mod } n & \quad \text{if $i$ is even} \\
-      g(i) \text{ mod }  n & \quad \text{if $i$ is odd}
+      g(i - 1) \pmod{n} & \quad \text{if $i$ is even} \\
+      g(i) \pmod{n} & \quad \text{if $i$ is odd}
     \end{cases}
@@
   \label{right-rotate}
-  a \rightturn i = \left[ a_{n - 1 + i\text{ mod }n} \ldots a_{i\text{ mod }n} \right]_{\left[ 2 \right]}, \text{ where } a = \left[ a_{n-1} \ldots a_0\right]_{\left[ 2 \right]} 
+  a \rightturn i = \left[ a_{{n - 1 + i}\pmod{n}} \ldots a_{i\pmod{n}} \right]_{\left[ 2 \right]}, \text{ where } a = \left[ a_{n-1} \ldots a_0\right]_{\left[ 2 \right]} 
 \end{equation}
@@
       0 & \quad \text{if $i = 0$}\\
-      e(i - 1) \oplus 2^{d(i-1)} \oplus 2^{g(i-1)}& \quad \text{otherwise} \\
+      gc(2 * \lfloor x-1/2 \rfloor) & \quad \text{otherwise} \\
     \end{cases}
@@
 \begin{algorithm}
-  \caption{The algorithm for calculating the Gray Code Inverse.     Given $g \in \mathbb{N}$, calculates the $i \in \mathbb{N}$ such that $gc\left( i \right) = g $ }
+  \caption{The algorithm for calculating the Gray Code Inverse of $g \in \mathbb{N}$, calculates the $i \in \mathbb{N}$ such that $gc\left( i \right) = g $ }
   \label{gray-code-inverse}
@@
     \State $ e \leftarrow e \oplus \left( e \left( w \right) \leftturn \left( d+ 1 \right) \right) $ 
-    \State $ d \leftarrow d + d \left( w \right) + 1 \text{ mod }n$ 
+    \State $ d \leftarrow d + d \left( w \right) + 1 \pmod{n}$ 
     \State $ h \leftarrow \left( h \ll n \right) | w $

### 2013-12-07 src/main.py
IT FUCKING WORKS

@@ -39,7 +39,11 @@ def set_knn(point, k, S):
     smallest_indices = k_smallest_indices(k, distances)
 
     # Find the points corresponding to those indices
-    nearest_neighbors = [S[i] for i in smallest_indices]
+    nearest_neighbors = []
+    for i in smallest_indices:
+        if S[i] != point:
+            nearest_neighbors.append(S[i])
+
     return nearest_neighbors
 
 def box_dist(point, box_center, box_radius):
@@ -52,7 +56,7 @@ def set_radius(S):
 
 def csearch(point, A_i, i, k, low, hi):
     if (hi-low) < V_CONSTANT:
-        return set_knn(point[i], k, list(set(A_i + point[low:hi])))
+        return set_knn(point[i], k, list(set(A_i + point[low:hi+1])))
 
     mid = int((hi + low) / 2)
     A_i = set_knn(point[i], k, list(set(A_i + [point[mid]])))
@@ -80,20 +84,20 @@ def construct(points, i, k):
     #print "point_i:", points[i]
     #print "A_i:", A_i
     # if p_i^ceil(rad(A_i)) < p_i+k
-    if points[i].shift(math.ceil(set_radius(A_i))) < points[min(i+k, len(points)-1)]:
+    if points[i].shift(math.ceil(set_radius(A_i))) < points[min(len(points)-1, i+k)]:
         upper = i
     else:
         I = 0
-        while points[i].shift(math.ceil(set_radius(A_i))) < points[min(len(points)-1, i+2**I)]:
+        while points[i].shift(math.ceil(set_radius(A_i))) < points[i+2**I-1]:
             I += 1
         upper = min(i + 2**I, len(points)-1)
 
     # if p_i^-ceil(rad(A_i)) > p_i-k
-    if points[i].shift(-1 * math.ceil(set_radius(A_i))) > points[max(0, i-k)]:
+    if points[i].shift(-1 * math.ceil(set_radius(A_i))) > points[i-k]:
         lower = i
     else:
         I = 0
-        while points[i].shift(-1 * math.ceil(set_radius(A_i))) > points[max(0, i-2**I)]:
+        while points[i].shift(-1 * math.ceil(set_radius(A_i))) > points[i-2**I]:
             I += 1
         lower = max(i - 2**I, 0)
 
@@ -108,6 +112,7 @@ def main(file_name, k):
[…]

### 2013-12-07 src/point.py

@@ -37,10 +37,7 @@ class Point(object):
         return morton.cmp_zorder(self.coordinates, other.coordinates) < 0
 
     def __eq__(self, other):
-        self.check_valid()
-        if self.sftype == 'hilbert':
-            return self.index_hilbert() == other.index_hilbert()
-        return morton.cmp_zorder(self.coordinates, other.coordinates) == 0
+        return self.coordinates == other.coordinates
 
     def __pow__(self, power):
         result = Point(tuple([x ** power for x in self.coordinates]), precision=self.precision, sftype=self.sftype)
@@ -57,3 +54,6 @@ class Point(object):
 
     def __repr__(self):
         return str(self.coordinates)
+
+def cmp_hilbert(a, b):
+    return a.index_hilbert() - b.index_hilbert()

### 2013-12-07 src/tester.py

@@ -2,11 +2,17 @@ from point import *
 import unittest
 
 class PointTests(unittest.TestCase):
+    a = Point((1,1,1))
+    b = Point((1,1,2))
+    
     def create_invalid(self):
         return Point((0,0,0), sftype='invalid')
     
     def test_create_invalid(self):
         self.assertRaises(ValueError, self.create_invalid)
+
+    def test_not_equal(self):
+        self.assertEqual(self.a != self.b, True)
 
 class HilbertPointTests(unittest.TestCase):
     a = Point((1,0,1), sftype='hilbert')

### 2013-12-07 src/main.py
Hilbert AKNN works

@@ -47,7 +47,7 @@ def set_knn(point, k, S):
     return nearest_neighbors
 
 def box_dist(point, box_center, box_radius):
-    return distance(point, box_center) - box_radius # This is roughly incorrect... Although it's a circle. So yeah.
+    return distance(point, box_center) - box_radius # This is correct... Although it's a circle. So yeah.
 
 # The radius of a set is the largest distance between any two points in the set S
 def set_radius(S):
@@ -108,12 +108,18 @@ def construct(points, i, k):
         
 
 def main(file_name, k):
-    dataset = [(x,y) for x in range(20) for y in range(20)]
-    #dataset = np.genfromtxt(str(file_name), delimiter="\t", dtype=np.uint32)
+    #dataset = [(x,y) for x in range(20) for y in range(20)]
+    dataset = np.genfromtxt(str(file_name), delimiter="\t", dtype=np.uint32)
 
+    print "Generating Hilbert points..."
     HP = [Point(item, sftype='hilbert') for item in dataset]
+    print "Sorting..."
     HP.sort(cmp=cmp_hilbert)
+
+    print "Generating Morton points..."
     MP = [Point(item, sftype='morton') for item in dataset]
+    print "Sorting..."
+    MP.sort(cmp=cmp_zorder)
 
     t_cum_aknn = 0
     t_cum_knn = 0
@@ -170,7 +176,7 @@ def main(file_name, k):
         t_cum_knn += tf_knn
         
         average_distance_from_opt += set_radius(AKNN_i) / set_radius(KNN_i)
-        print HP[i], "\tAKNN:", AKNN_i, "\tOPT:", KNN_i
+        #print HP[i], "\tAKNN:", AKNN_i, "\tOPT:", KNN_i
     print ""
     print "Time Hilbert AKNN:", t_cum_aknn
     print "Time OPT:", t_cum_knn

### 2013-12-07 src/point.py

@@ -57,3 +57,6 @@ class Point(object):
 
 def cmp_hilbert(a, b):
     return a.index_hilbert() - b.index_hilbert()
+
+def cmp_zorder(a, b):
+    return morton.cmp_zorder(a.coordinates, b.coordinates)

### 2013-12-07 src/main.py
IT WORKS AMAZZZZZINGLYYYYY

@@ -121,42 +121,17 @@ def main(file_name, k):
     print "Sorting..."
     MP.sort(cmp=cmp_zorder)
 
-    t_cum_aknn = 0
-    t_cum_knn = 0
+
     k = int(k)
     precision = 32
-    """
-    print "Running Z-order AKNN comparison on", len(P), "points..."
-    average_distance_from_opt = 0
-    for i in range(len(P)):
-        iter_fraction = float(i) / len(P)
-
-        # Progress bar
-        amtDone = iter_fraction * 100
-        sys.stdout.write("\r%.1f%%" %amtDone)
-        sys.stdout.flush()
-
-        t0_aknn = time.time()
-        AKNN_i = construct_morton(P, i, k)
-        tf_aknn = time.time() - t0_aknn
-        t_cum_aknn += tf_aknn
-
-        t0_knn = time.time()
-        KNN_i = set_knn(P[i], k, P)
-        tf_knn = time.time() - t0_knn
-        t_cum_knn += tf_knn
-        
-        average_distance_from_opt += set_radius(AKNN_i) / set_radius(KNN_i)
+    t_cum_hilbert = 0
+    t_cum_morton = 0
+    t_cum_naive = 0
+    
     print ""
-    print "Time Morton AKNN:", t_cum_aknn
-    print "Time OPT:", t_cum_knn
-    print "Average times distance from OPT:", average_distance_from_opt / len(P)
-    """
-    t_cum_aknn = 0
-    t_cum_knn = 0
-    print ""
-    print "Running Hilbert AKNN comparison on", len(HP), "points..."
-    average_distance_from_opt = 0
+    print "Running AKNN comparison on", len(HP), "points in", len(HP[0]),"dimensions..."
+    distance_ratio_hilbert = 0
+    distance_ratio_morton = 0
     for i in range(len(HP)):
         iter_fraction = float(i) / len(HP)
 
@@ -165,22 +140,31 @@ def main(file_name, k):
         sys.stdout.write("\r%.1f%%" %amtDone)
         sys.stdout.flush()
 
-        t0_aknn = time.time()
-        AKNN_i = construct(HP, i, k)
-        tf_aknn = time.time() - t0_aknn
-        t_cum_aknn += tf_aknn
+        t0_hilbert = time.time()
+        AKNN_hilbert = construct(HP, i, k)
+        tf_hilbert = time.time() - t0_hilbert
+        t_cum_hilbert += tf_hilbert
 
[…]

### 2013-12-07 src/point.py

@@ -30,6 +30,9 @@ class Point(object):
         result = Point(tuple([x + amount for x in self.coordinates]), precision=self.precision, sftype=self.sftype)
         return result
     
+    def __len__(self):
+        return self.dimension
+
     def __lt__(self, other):
         self.check_valid()
         if self.sftype == 'hilbert':

### 2013-12-07 src/main.py
Added z-order aknn

@@ -129,7 +129,7 @@ def main(file_name, k):
     t_cum_naive = 0
     
     print ""
-    print "Running AKNN comparison on", len(HP), "points in", len(HP[0]),"dimensions..."
+    print "Running AKNN comparison on", len(HP), "points in", len(HP[0]), "dimensions..."
     distance_ratio_hilbert = 0
     distance_ratio_morton = 0
     for i in range(len(HP)):

### 2013-12-07 src/main.py
Added correct attribution to all files

@@ -4,18 +4,19 @@
 # space and find k-nearest neighbors based on this ordering
 
 # CPSC 4110 - Approximation Algorithms
-# Camara Lerner & Chris Rabl
+# Implemented by:
+# Camara Lerner - https://github.com/lernerc
+# Christopher Rabl - https://github.com/crabl 
 
-import numpy
+# Based on algorithm detailed by Michael Connor and Piyush Kumar
+# http://doi.ieeecomputersociety.org/10.1109/TVCG.2010.9
+
+
+import numpy as np
 import sys
 import math
 import bitstring
-import numpy as np 
 import time
-import warnings
-
-# Ignore exponent divide-by-zero warning
-#warnings.simplefilter("ignore", RuntimeWarning) 
 
 from point import *
 
@@ -27,8 +28,7 @@ def distance(a, b):
 
 # Find the indices of the k smallest items in a set S, assuming that S contains a zero element
 def k_smallest_indices(k, S):
-    # We have to look at [1:k+1] because we assume that S contains the point
-    # that we are looking for
+    # We have to look at [1:k+1] because we assume that S contains the point that we are looking for
     return [idx for (s, idx) in sorted(zip(S,range(len(S))))][1:k+1]
 
 # Find the k nearest neighbors to a point in a given set S
@@ -52,7 +52,7 @@ def box_dist(point, box_center, box_radius):
 # The radius of a set is the largest distance between any two points in the set S
 def set_radius(S):
     distances = [distance(p0, p1) for p0 in S for p1 in S]
-    return np.max(distances)
+    return max(distances)
 
 def csearch(point, A_i, i, k, low, hi):
     if (hi-low) < V_CONSTANT:
@@ -64,11 +64,11 @@ def csearch(point, A_i, i, k, low, hi):
         return A_i
     if point[i] < point[mid]:
         A_i = csearch(point, A_i, i, k, low, mid - 1)
-        if point[mid] < point[i].shift(np.ceil(set_radius(A_i))):
+        if point[mid] < point[i].shift(math.ceil(set_radius(A_i))):
             A_i = csearch(point, A_i, i, k, mid + 1, hi)
     else:
         A_i = csearch(point, A_i, i, k, mid + 1, hi)
-        if point[i].shift(-1 * np.ceil(set_radius(A_i))) < point[mid]:
[…]

### 2013-12-07 src/point.py

@@ -1,3 +1,7 @@
+# Multidimensional Point Class
+# Supports Hilbert and Z-Order Comparison
+# Copyright (C) 2013 Christopher Rabl
+
 import math
 
 import hilbert

### 2013-12-07 src/draw_hilbert.py
...just kidding, now they're all done

@@ -1,7 +1,10 @@
+# Hilbert Curve Visualization Code
+# Copyright (C) 2013 Christopher Rabl
+
 import numpy as np
 import matplotlib.pyplot as plt
 from mpl_toolkits.mplot3d import Axes3D
-from hilbert import *
+from hilbert_old import *
 import random
 
 def draw_graph_2d(points, path, file_name):

### 2013-12-07 src/draw_morton.py

@@ -1,3 +1,6 @@
+# Z-Order Curve Visualization Code
+# Copyright (C) 2013 Christopher Rabl
+
 import numpy as np
 import matplotlib.pyplot as plt
 from mpl_toolkits.mplot3d import Axes3D

### 2013-12-07 src/tester.py

@@ -1,10 +1,13 @@
+# Application Unit Tests
+# Copyright (C) 2013 Christopher Rabl
+
 from point import *
 import unittest
 
 class PointTests(unittest.TestCase):
     a = Point((1,1,1))
     b = Point((1,1,2))
-    
+
     def create_invalid(self):
         return Point((0,0,0), sftype='invalid')
     

### 2013-12-07 src/utils.py

@@ -1,3 +1,7 @@
+# Copyright (C) 2013 Aldo Cortesi
+# Adapted from source code by Aldo Cortesi
+# https://github.com/cortesi/scurve
+
 """
     A lot of these functions are written for clarity rather than speed. We'll
     fix that in time.

### 2013-12-07 paper/paper.tex
added more to the paper

@@
 \author{Camara Lerner \hspace{2cm} Christopher Rabl \\
-  Department of Mathematics and Computer Science\\
+  Department of Mathematics and Computer Science \\
   University of Lethbridge, Canada}
@@
 
-\begin{abstract}
+\section{Introduction}
 
-\end{abstract}
-
-\section{Introduction}
+\section{Morton Curve}
 
@@
 The space that the curve fills can be in any dimensionally space $(n)$ with the length of each dimension filled equal to $2^m$.
-This means that for each point in the curve the maximum bit size for each dimension is $(m)$~\cite{Hamilton:2006}.
+This means that for each point in the curve the maximum bit size for each dimension is $(m)$.
+The algorithm used is adapted from~\cite{Hamilton:2006}.
 
 The curve labels the points from $0$ to $2^{m*n} - 1$ in the order that the curve visits each point.
-The Hilbert Curve considered here will start in the bottom left hand corner and finishes in the upper left corner~\cite{Hamilton:2006}.
+The Hilbert Curve considered here will start in the bottom left hand corner and finishes in the upper left corner.
 
@@
 These facts are used to inverse the operation of and get the point from the index in the list for both gray code and Hilbert indexing.
-Hilbert curves also can use information from gray code in determining the changes to the unit Hilbert curve~\cite{Hamilton:2006}.
+Hilbert curves also can use information from gray code in determining the changes to the unit Hilbert curve.
 
 \flushleft
-Formula for Gray Code is in \ref{gray-code}.
+Formula for Gray Code is in~\ref{gray-code}.
 \begin{equation}
@@
 
-Formula for $g$ is displayed in \ref{g}.
+\begin{algorithm}
+  \caption{The algorithm for calculating the Gray Code Inverse of $g \in \mathbb{N}$, calculates the $i \in \mathbb{N}$ such that $gc\left( i \right) = g $ }
+  \label{gray-code-inverse}
+  \begin{algorithmic}[1]
+    \Require $g \in \mathbb{N}$ 
+    \State $ m \leftarrow \text{ number of bits to represent } g$ 
+    \State $ \left( i, j \right) \leftarrow \left( g, 1 \right) $ 
+    \While{$ j < m $ }
+    \State $ i \leftarrow i \oplus \left( g \gg j \right)$ 
+    \State $ j \leftarrow j + 1$ 
+    \EndWhile \\
+    \Return $i$ % \in \mathbb{N} \text{ such that } gc \left( i \right) = g $
+  \end{algorithmic}
+\end{algorithm}
+
+From one point in the Hilbert curve to another since only one bit changes by onepoint to another there is a formula to determine which bit changes.
+The formula for this returns the bit that will change next~\ref{g}.
 \begin{equation}
@@
 
-Formula for $d(i)$ is displayed in \ref{d}.
+The direction of the unit Hilbert curve is represented by~\ref{d}.
 \begin{equation}
@@
 
-Formula for the bitwise right rotation is displayed in \ref{right-rotate}. The bitwise left rotation is similar.
-\begin{equation}
-  \label{right-rotate}
-  a \rightturn i = \left[ a_{{n - 1 + i}\pmod{n}} \ldots a_{i\pmod{n}} \right]_{\left[ 2 \right]}, \text{ where } a = \left[ a_{n-1} \ldots a_0\right]_{\left[ 2 \right]} 
-\end{equation}
-
-Formula for $e$ is displayed in \ref{e}.
+The entry point of the unit Hilbert curve is represented by~\ref{e}.
 \begin{equation}
@@
 
-\begin{algorithm}
-  \caption{The algorithm for calculating the Gray Code Inverse of $g \in \mathbb{N}$, calculates the $i \in \mathbb{N}$ such that $gc\left( i \right) = g $ }
-  \label{gray-code-inverse}
-  \begin{algorithmic}[1]
-    \Require $g \in \mathbb{N}$
-    \Return $i \in \mathbb{N} \text{ such that } gc \left( i \right) = g $
-    \State $ m \leftarrow \text{ number of bits to represent } g$ 
-    \State $ \left( i, j \right) \leftarrow \left( g, 1 \right) $ 
-    \While{$ j < m $ }
-    \State $ i \leftarrow i \oplus \left( g \gg j \right)$ 
-    \State $ j \leftarrow j + 1$ 
-    \EndWhile
-    % \oplus is exclusive or
-    % \ll left shift
-    % \gg right shift
-    % \mid bitwise or
-  \end{algorithmic}
-\end{algorithm}
+Formula for the bitwise right rotation is displayed in \ref{right-rotate}. The bitwise left rotation is similar.
+\begin{equation}
+  \label{right-rotate}
+  a \rightturn i = \left[ a_{{n - 1 + i}\pmod{n}} \ldots a_{i\pmod{n}} \right]_{\left[ 2 \right]}, \text{ where } a = \left[ a_{n-1} \ldots a_0\right]_{\left[ 2 \right]} 
+\end{equation}
+
+In determining the Hilbert curve index the $l$ variable used in~\ref{hilbert-point-to-index} has each bit representing whether the point point is above or below the axis corresponding to each bit.
+And the $e$ and $d$ variables represent the rotation and reflections nessesary to map the current sub-Hilbert curve containing the point ${\bf p}$ to the normal orientation of sub-Hilbert curve.
+These values are recalculated until the sub-Hilbert curve is a unit Hilbert curve.
+There is also a variable $w$, which is calculated at every iteration that is ORed wit$h$ after $h$ is bit-shifted left one bit.
+Then after all the iterations $h$ is the index of the point ${\bf p}$.
+The algorithn runs in $O(m^2)$ time.
 
 \begin{algorithm}
-  \caption{Calculating the Hilbert index given any dimensional point ${\bf p}$ 
+  \caption{Calculates the Hilbert index given any dimensional point ${\bf p}$ 
     of size $n$ as long as the bits used within an index of the hilbert curve 
@@
     \Require $n, m \in \mathbb{N} - \{0\} \text{ and a point } {\bf p} \in \mathbb{N}^n$ 
-    \Return $h \in \mathbb{N} \text{, the Hilbert index of the point }{\bf p}$
     \State $ \left( h, e, d \right) \leftarrow \left( 0, 0, 0 \right) $
@@
     \State $ l \leftarrow \left[ \left( p_{n-1} , i \right) \ldots \left( p_0 , i \right) \right]_{\left[ 2 \right]} $ 
-    \State $ l \leftarrow \left( b \oplus e \right) \rightturn \left( d+1 \right)$ 
+    \State $ l \leftarrow \left( l \oplus e \right) \rightturn \left( d+1 \right)$ 
     \State $ w \leftarrow gc^{-1} \left( l \right)$ 
@@
     \State $ h \leftarrow \left( h \ll n \right) | w $
-    \EndFor
+    \EndFor \\
+    \Return $h $%\in \mathbb{N} \text{, the Hilbert index of the point }{\bf p}$
+
   \end{algorithmic}
 \end{algorithm}
+
+\section{Algorithm}
+
+\section{Results}
 

### 2013-12-07 src/gentestdata.py
Added data generator

@@ -1,0 +1,8 @@
+f = open("lattice_40x40.csv", "w")
+
+points = [(x,y) for x in range(40) for y in range(40)]
+
+for (x,y) in points:
+    f.write(str(x)+"\t"+str(y)+"\n")
+
+f.close()

### 2013-12-07 paper/paper.tex
fixed the labeling of gray code inverse

@@
 \begin{algorithm}
-  \caption{The algorithm for calculating the Gray Code Inverse of $g \in \mathbb{N}$, calculates the $i \in \mathbb{N}$ such that $gc\left( i \right) = g $ }
+  \caption{The algorithm for calculating the Gray Code Inverse $(gc^{-1}(g))$ of $g \in \mathbb{N}$, calculates the $i \in \mathbb{N}$ such that $gc\left( i \right) = g $ }
   \label{gray-code-inverse}

### 2013-12-07 paper/mkgraph.cc
added the mkgraph template for you

@@ -1,0 +1,161 @@
+#include <iostream>
+#include <fstream>
+#include <string>
+using namespace std;
+
+const int IMAGE_NUM = 5;
+const string imagesets[IMAGE_NUM] = {"combo", "galwaycity", "goes",
+				     "joetouristweather", "ultrapig2004"};
+const string shortImageSets[IMAGE_NUM] = {"combo", "galway", "goes", "joe",
+					  "ultrapig"};
+const int WEIGHT_NUM = 4;
+const string weights[WEIGHT_NUM] = {"RMSE", "SSIM", "NQM error", "DCTune"};
+const char shortweights[WEIGHT_NUM] = {'r', 's', 'n', 'd'};
+
+const int METHOD_NUM = 4;
+const string methods[METHOD_NUM] = {"mst", "msta", "centroid", "traditional"};
+
+const int COMPRESSION_NUM = 2;
+const char compression[COMPRESSION_NUM] = {'j', 'w'};
+
+// This file takes "imageset weight method compression" and then creates the gnuplot text file that graphs the lagrange distortion curve compression numbers, and the same size compression numbers and if available the lagrange closest sizes compression
+
+int main(int argv, char *argc[]) {
+   if(argv < 5) {
+      cout << "not enough arguments" << endl;
+      cout << "imageset weight method compression" << endl;
+   }
+   
+   int name = -1, w = -1, m = -1, c = -1;
+   // checking name
+   for(int i = 0; i < IMAGE_NUM; i++) {
+      if(shortImageSets[i] == string(argc[1])) {
+	 name = i;
+      }
+   }
+   if(name == -1) {
+      cerr << "Not a proper image set" << endl;
+      for(int i = 0; i < IMAGE_NUM; i++) {
+	 cerr << shortImageSets[i] << " ";
+      }
+      cerr << endl;
+      return 1;
+   }
+   // checking weight
+   for(int i = 0; i < WEIGHT_NUM; i++) {
+      if(shortweights[i] == argc[2][0]) {
+	 w = i;
+      }
+   }
+   if(w == -1) {
+      cerr << "Not a proper weight" << endl;
+      for(int i = 0; i < WEIGHT_NUM; i++) {
+	 cerr << shortweights[i] << " ";
+      }
+      cerr << endl;
+      return 1;
+   }
+   // checking method
+   for(int i = 0; i < METHOD_NUM; i++) {
[…]

### 2013-12-07 src/gentestdata.py
Generating random test data now

@@ -1,8 +1,10 @@
-f = open("lattice_40x40.csv", "w")
+import random
 
-points = [(x,y) for x in range(40) for y in range(40)]
+f = open("lattice_40x40_random.csv", "w")
 
-for (x,y) in points:
-    f.write(str(x)+"\t"+str(y)+"\n")
+points = [(random.randint(0,100),random.randint(0,100),random.randint(0,100)) for x in range(40) for y in range(40) for  z in range(40)]
+
+for (x,y,z) in points:
+    f.write(str(x)+"\t"+str(y)+"\t"+str(z)+"\n")
 
 f.close()

### 2013-12-07 paper/paper.tex
added the new algorithms to the paper

@@
 \section{Morton Curve}
+
+\begin{algorithm}
+  \caption{COMPARE(point ${\bf p}$, point ${\bf q}$)}
+  \label{compare}
+  \begin{algorithmic}[1]
+    \State $x \leftarrow 0, dim \leftarrow 0$
+    \ForAll{j=0 to d}
+    \State $y \leftarrow \text{FirstMSB}(p_{[j]}, q_{[j]})$
+    \If{$x < y$}
+    \State $x \leftarrow y, dim \leftarrow j$
+    \EndIf 
+    \EndFor \\
+    \Return $p_{[dim]} < q_{[dim]}$
+  \end{algorithmic}
+\end{algorithm}
+
+\begin{algorithm}
+  \caption{FirstMSB(double $a$, double $b$)}
+  \label{first-msb}
+  \begin{algorithmic}[1]
+    \State $w \leftarrow x < y \text{ and } x < (x \mathrel{\&} y)$ \\
+    \Return $w$
+  \end{algorithmic}
+\end{algorithm}
+
+\begin{algorithm}
+  \caption{KNN Graph Construction Algorithm}
+  \label{knn-graph-alg}
+  \begin{algorithmic}[1]
+    \Require $\text{Mortan order compare operators: } <, > 
+    \text{ and Hilbert order compare operators: } <, >$
+    \Ensure $A_i \text{ contains } k \text{ nearest neighbors of } p_i \text{ in } P \text{.}$
+    \State $P \leftarrow \text{QuickSort}(P, <)$
+    \ForAll{$p_i$ in $P$}
+    \State $A_i \leftarrow \text{SET-KNN}(p_i, \{p_{i-k}, \dots, p_{i+k} \})$
+    \If{$\text{SHIFT}(p_i, \lceil \text{SET-RAD}(A_i) \rceil) < p_{i+k}$}
+    \State $upper \leftarrow i$
+    \Else
+    \State $I \leftarrow 1$
+    \While{$\text{SHIFT}(p_i, \lceil \text{SET-RAD}(A_i) \rceil) < p_{i+2^I}$}
+    \State $++I$
+    \EndWhile
+    \State $upper \leftarrow \text{min}(i+2^I, n)$
+    \EndIf
+    \If{$\text{SHIFT}(p_i, - \lceil \text{SET-RAD}(A_i) \rceil) > p_{i-k}$}
+    \State $lower \leftarrow i$
+    \Else
+    \State $I \leftarrow 1$
+    \While{$\text{SHIFT}(p_i, - \lceil \text{SET-RAD}(A_i) \rceil)$}
+    \State $++I$
+    \EndWhile
+    \State $lower \leftarrow \text{max}(i-2^I, 1)$
+    \EndIf
+    \If{$lower \not= upper$}
+    \State $A_i \leftarrow \text{CSEARCH}(p_i, lower, upper)$
+    \EndIf
+    \EndFor \\
+    \Return $A_i$
+  \end{algorithmic}
+\end{algorithm}
+
+\begin{algorithm}
+  \caption{CSEARCH(point $p_i$, int $lo$, int $hi$)}
+  \label{csearch}
+  \begin{algorithmic}[1]
+    \If{$(hi - lo) < v $}
+    \State $A_i \leftarrow \text{SET-KNN}(p_i, A_i \cup \{ p_{lo} \dots p_{hi} \})$
+    \Return $A_i$
+    \EndIf
+    \State $mid \leftarrow (hi+lo)/2$
+    \State $A_i \leftarrow \text{SET-KNN}(p_i, A_i \cup p_{mid})$
+    \If{$\text{DIST}(p_i, \text{CIRCLE}(p_{lo}, p_{hi})) \ge \text{SET-RAD}(A_i)$} 
+    \Return $A_i$
+    \EndIf
+    \If{$p_i < p_{mid}$}
+    \State $A_i \leftarrow \text{CSEARCH}(p_i, lo, mid-1)$
+    \If{$p_{mid} < \text{SHIFT}(p_i, \lceil \text{SET-RAD}(A_i) \rceil)$}
+    \State $A_i \leftarrow \text{CSEARCH}(p_i, mid+1, hi)$
+    \EndIf
+    \Else
+    \State $A_i \leftarrow \text{CSEARCH}(p_i, mid+1, hi)$
+    \If{$\text{SHIFT}(p_i, - \lceil \text{SET-RAD}(A_i) \rceil) < p_{mid}$}
+    \State $A_i \leftarrow \text{CSEARCH}(p_i, low, mid - 1)$
+    \EndIf
+    \EndIf \\
+    \Return $A_i$
+
+  \end{algorithmic}
+\end{algorithm}
 
@@
 From one point in the Hilbert curve to another since only one bit changes by onepoint to another there is a formula to determine which bit changes.
-The formula for this returns the bit that will change next~\ref{g}.
+The formula for this returns the bit that will change next~(\ref{g}).
 \begin{equation}
@@
 
-The direction of the unit Hilbert curve is represented by~\ref{d}.
+The direction of the unit Hilbert curve is represented by~(\ref{d}).
 \begin{equation}
@@
 
-The entry point of the unit Hilbert curve is represented by~\ref{e}.
+The entry point of the unit Hilbert curve is represented by~(\ref{e}).
 \begin{equation}
@@
 
-Formula for the bitwise right rotation is displayed in \ref{right-rotate}. The bitwise left rotation is similar.
+Formula for the bitwise right rotation is displayed in~(\ref{right-rotate}). The bitwise left rotation is similar.
 \begin{equation}
@@
 
-In determining the Hilbert curve index the $l$ variable used in~\ref{hilbert-point-to-index} has each bit representing whether the point point is above or below the axis corresponding to each bit.
+In determining the Hilbert curve index the $l$ variable used in~(\ref{hilbert-point-to-index}) has each bit representing whether the point point is above or below the axis corresponding to each bit.
 And the $e$ and $d$ variables represent the rotation and reflections nessesary to map the current sub-Hilbert curve containing the point ${\bf p}$ to the normal orientation of sub-Hilbert curve.

### 2013-12-07 src/gentestdata.py
Removed bug and added data files

@@ -2,9 +2,9 @@ import random
 
 f = open("lattice_40x40_random.csv", "w")
 
-points = [(random.randint(0,100),random.randint(0,100),random.randint(0,100)) for x in range(40) for y in range(40) for  z in range(40)]
+points = [(random.randint(0,100),random.randint(0,100)) for x in range(40) for y in range(40)]
 
-for (x,y,z) in points:
-    f.write(str(x)+"\t"+str(y)+"\t"+str(z)+"\n")
+for (x,y) in points:
+    f.write(str(x)+"\t"+str(y)+"\n")
 
 f.close()

### 2013-12-07 src/main.py

@@ -11,7 +11,6 @@
 # Based on algorithm detailed by Michael Connor and Piyush Kumar
 # http://doi.ieeecomputersociety.org/10.1109/TVCG.2010.9
 
-
 import numpy as np
 import sys
 import math
@@ -102,16 +101,16 @@ def construct(points, i, k):
     return A_i
         
 
-def main(file_name, k):
+def main(file_name, k, validate):
     #dataset = [(x,y) for x in range(20) for y in range(20)]
     dataset = np.genfromtxt(str(file_name), delimiter="\t", dtype=np.uint32)
 
-    print "Generating Hilbert points:",
+    print "Generating Hilbert ordering:",
     HP = [Point(item, sftype='hilbert') for item in dataset]
     print "Sorting..."
-    HP.sort(key=lambda h: h.index_hilbert())
+    HP.sort(key=lambda h: h.index_hilbert()) # Have to use this or we run into issues with 'long'
 
-    print "Generating Morton points:",
+    print "Generating Morton ordering:",
     MP = [Point(item, sftype='morton') for item in dataset]
     print "Sorting..."
     MP.sort(cmp=cmp_zorder)
@@ -121,8 +120,9 @@ def main(file_name, k):
     precision = 32
     t_cum_hilbert = 0
     t_cum_morton = 0
-    t_cum_naive = 0
-    
+    t_cum_naive0 = 0
+    t_cum_naive1 = 0
+
     print ""
     print "Approximate", str(k)+"-Nearest Neighbours"
     print "Running AKNN comparison on", len(HP), "points in", len(HP[0]), "dimensions..."
@@ -133,8 +133,9 @@ def main(file_name, k):
 
         # Progress bar
         amtDone = iter_fraction * 100
-        sys.stdout.write("\r%.1f%%" %amtDone)
-        sys.stdout.flush()
+        if not validate:
+            sys.stdout.write("\r%.1f%%" %amtDone)
+            sys.stdout.flush()
 
         t0_hilbert = time.time()
         AKNN_hilbert = construct(HP, i, k)
@@ -146,21 +147,42 @@ def main(file_name, k):
         tf_morton = time.time() - t0_morton
         t_cum_morton += tf_morton
 
-        t0_naive = time.time()
-        KNN_i = set_knn(HP[i], k, HP)
-        tf_naive = time.time() - t0_naive
-        t_cum_naive += tf_naive
+        t0_naive0 = time.time()
[…]

### 2013-12-07 src/point.py

@@ -13,12 +13,14 @@ class Point(object):
         self.dimension = len(coordinates)
         self.precision = precision
         self.sftype = sftype.lower()
+        # Ensure that we have a space-filling curve type defined that we support,
+        # if not, go ahead and write it! Just be sure to add it in here ;)
         if not self.sftype in ['hilbert', 'morton']:
             raise ValueError('Point type must be either Hilbert or Morton')
         
     def check_valid(self):
+        # Check whether any coordinate vertex is negative
         if any([x < 0 for x in self.coordinates]):
-            #raise ValueError('Cannot get Hilbert index of negative point')
             return False
         return True
 
@@ -37,6 +39,8 @@ class Point(object):
     def __len__(self):
         return self.dimension
 
+    # Check whether a point's order is less than another point's order using
+    # the appropriate comparison operator
     def __lt__(self, other):
         self.check_valid()
         if self.sftype == 'hilbert':
@@ -46,13 +50,16 @@ class Point(object):
     def __eq__(self, other):
         return self.coordinates == other.coordinates
 
+    # Used in distance calculation
     def __pow__(self, power):
         result = Point(tuple([x ** power for x in self.coordinates]), precision=self.precision, sftype=self.sftype)
         return result
 
+    # Should really use this in SHIFT, but I am too lazy
     def __add__(self, other):
         return Point(tuple([self[i] - other[i] for i in range(self.dimension)]), precision=self.precision, sftype=self.sftype)
     
+    # Used in distance calculation
     def __sub__(self, other):
         return Point(tuple([self[i] - other[i] for i in range(self.dimension)]), precision=self.precision, sftype=self.sftype)
 
@@ -62,8 +69,10 @@ class Point(object):
     def __repr__(self):
         return str(self.coordinates)
 
+# Hilbert order comparison operator, used in QuickSort
 def cmp_hilbert(a, b):
     return a.index_hilbert() - b.index_hilbert()
 
+# Morton order comparison operator, used in QuickSort
 def cmp_zorder(a, b):
     return morton.cmp_zorder(a.coordinates, b.coordinates)

### 2013-12-07 src/tester.py

@@ -1,8 +1,10 @@
 # Application Unit Tests
 # Copyright (C) 2013 Christopher Rabl
 
-from point import *
 import unittest
+
+from point import *
+from main import *
 
 class PointTests(unittest.TestCase):
     a = Point((1,1,1))
@@ -38,6 +40,78 @@ class MortonPointTests(unittest.TestCase):
         self.assertEqual(self.a < self.b, True)
         self.assertEqual(self.a == self.b, False)
         self.assertEqual(self.a > self.b, False)
+
+class AlgorithmCorrectnessTests(unittest.TestCase):
+    dataset = [(x,y) for x in range(4) for y in range(4)]
+
+    # Generate and sort Hilbert points
+    HP = [Point(item, sftype='hilbert') for item in dataset]
+    HP.sort(key=lambda h: h.index_hilbert()) # Have to use this or we run into issues with 'long'
+
+    # Generate and sort Morton points
+    MP = [Point(item, sftype='morton') for item in dataset]
+    MP.sort(cmp=cmp_zorder)
+
+    k = 4
+
+    def test_hilbert_construct(self):
+        # First case will only return us k-1 results because it is the bottom corner
+        correct = [[(0, 1), (1, 0), (1, 1)], 
+        [(0, 0), (1, 1), (1, 0), (2, 0)],
+        [(0, 1), (1, 0), (2, 0), (0, 0)],
+        [(2, 0), (0, 0), (1, 1), (0, 1)],
+        [(1, 0), (2, 1), (3, 0), (1, 1)],
+        [(2, 0), (3, 1), (2, 1), (1, 0)],
+        [(3, 0), (2, 1), (3, 2), (2, 0)],
+        [(2, 0), (3, 1), (2, 2), (1, 0)],
+        [(3, 2), (2, 1), (2, 3), (3, 1)],
+        [(2, 2), (3, 1), (3, 3), (2, 1)],
+        [(3, 2), (2, 3), (2, 2), (3, 1)],
+        [(2, 2), (3, 3), (1, 3), (3, 2)],
+        [(1, 2), (2, 3), (0, 3), (2, 2)],
+        [(0, 2), (1, 3), (0, 3), (2, 3)],
+        [(1, 2), (0, 3), (1, 3), (2, 3)],
+        [(0, 2), (1, 3), (1, 2), (2, 3)]]
+
+        valid = []
+        for arr in correct:
+            temp_arr = []
+            for p in arr:
[…]

### 2013-12-07 src/tester.py
Unit tests should work

@@ -88,7 +88,7 @@ class AlgorithmCorrectnessTests(unittest.TestCase):
         correct = [[(0, 1), (1, 0), (1, 1)],
         [(1, 1), (0, 0), (0, 2), (1, 0)],
         [(1, 1), (0, 0), (0, 1), (0, 2)],
-        [(0, 1), (1, 0), (1, 2), (0, 0)],
+        [(0, 1), (1, 0), (1, 2), (0, 2)],
         [(0, 1), (1, 2), (0, 3), (1, 1)],
         [(0, 2), (1, 3), (1, 2), (0, 1)],
         [(1, 1), (0, 2), (1, 3), (0, 3)],
@@ -96,7 +96,7 @@ class AlgorithmCorrectnessTests(unittest.TestCase):
         [(2, 1), (3, 0), (3, 1), (1, 2)],
         [(2, 0), (3, 1), (2, 2), (3, 0)],
         [(2, 0), (3, 1), (2, 1), (2, 2)],
-        [(3, 2), (2, 1), (3, 0), (2, 0)],
+        [(3, 2), (2, 1), (3, 0), (2, 2)],
         [(3, 2), (2, 1), (2, 3), (3, 1)],
         [(2, 2), (3, 3), (3, 2), (2, 1)],
         [(2, 2), (3, 1), (3, 3), (2, 3)],
@@ -108,7 +108,7 @@ class AlgorithmCorrectnessTests(unittest.TestCase):
                 temp_arr.append(Point(p, sftype='morton'))
             valid.append(temp_arr)
 
-        for i in range(len(self.HP)):
+        for i in range(len(self.MP)):
             self.assertEqual(all([p in valid[i] for p in construct(self.MP, i, self.k)]), True)
 
 

### 2013-12-07 paper/paper.tex
added the graphs to the paper

@@
 
+\begin{figure}
+\begin{center}
+\includegraphics[scale=0.5]{YeaGra.png}
+\caption{The Runtime Analysis of the Yeast Database.}
+\label{run-yeast}
+\end{center}
+\end{figure}
+
+\begin{figure}
+\begin{center}
+\includegraphics[scale=0.5]{ArcGra.png}
+\caption{The Runtime Analysis of the Acrene Database.}
+\label{run-arcene}
+\end{center}
+\end{figure}
+
+\begin{figure}
+\begin{center}
+\includegraphics[scale=0.5]{LatGra.png}
+\caption{The Runtime Analysis of the Lattice Database.}
+\label{run-lattice}
+\end{center}
+\end{figure}
+
+\begin{figure}
+\begin{center}
+\includegraphics[scale=0.5]{YeaGra2.png}
+\caption{The Approximation bound of the Yeast Database.}
+\label{approx-yeast}
+\end{center}
+\end{figure}
+
+\begin{figure}
+\begin{center}
+\includegraphics[scale=0.5]{ArcGra2.png}
+\caption{The Approximation bound of the Acrene Database.}
+\label{approx-acrene}
+\end{center}
+\end{figure}
+
+\begin{figure}
+\begin{center}
+\includegraphics[scale=0.5]{LatGra2.png}
+\caption{The Approximation bound of the Lattice Database.}
+\label{approx-lattice}
+\end{center}
+\end{figure}
+
 \bibliographystyle{amsplain}

### 2013-12-07 paper/paper.tex
Added sections to the paper, Introduction, Correctness, First Method

@@
 \section{Introduction}
+
+All methods of calculating K-Nearest Neighbours (KNN) are marred by the ``curse of dimensionality'' \cite{Bellman:2003}.
+Even though it is not an NP-Complete problem, it is still difficult to calculate KNN in high dimensions.
+Approaches such as spatially ordered trees (such as quadtrees and $kd$-trees) and space-filling curves have been applied to this problem in order to solve it more quickly than the naive approach (calculating a distance matrix between each point and choosing the smallest $k$ entries from the row corresponding to the query point $q$).
+
+\section{Method of Dahl and Wootters}
+
+Contrary to the project proposal, the authors decided against exploring the approach of Dahl and Wootters \cite{Dahl:2008} due to a significant lack of implementation specificity: many important details were completely omitted from the paper and could not be extrapolated based on the context of the statements in which they appeared.
+Instead, a variant of method by Connor and Kumar \cite{Connor:2010} was chosen, since comparing and contrasting the neighbourhood-preserving properties of each space-filling curve turned out to be a more fruitful and unique comparison.
+
+\section{Space-Filling Curves}
 
@@
   \begin{algorithmic}[1]
-    \Require $\text{Mortan order compare operators: } <, > 
-    \text{ and Hilbert order compare operators: } <, >$
+    \Require $\text{Morton order comparison operator: } < 
+    \text{ and Hilbert order compare operator: } <$
     \Ensure $A_i \text{ contains } k \text{ nearest neighbors of } p_i \text{ in } P \text{.}$
@@
 
-The Hilbert Curve is a space filling curve that makes the next point in the curve only one unit distance from the previous point.
+The Hilbert Curve is a space-filling curve that is based on a generalization of Gray codes in $d$ dimensions.
 The curve is defined based on the amount of space the curve fills.
@@
 
-\section{Algorithm}
+\section{Implementation Correctness}
+
+In order to ensure the correctness of our implementation, some small test cases were generated and verified by hand over the course of the development process.
+These have been digitized and implemented in the form of Python unit tests using the UnitTest library.
+Additionally, the correctness of the ordering operators were verified using a similar approach.
 
 \section{Results}
+
+More text here
 
@@
 \includegraphics[scale=0.5]{YeaGra.png}
-\caption{The Runtime Analysis of the Yeast Database.}
+\caption{Running time analysis of the Yeast dataset ($d = 8$, $n = 1484$).}
 \label{run-yeast}
@@
 \includegraphics[scale=0.5]{ArcGra.png}
-\caption{The Runtime Analysis of the Acrene Database.}
+\caption{Running time analysis of the ARCENE dataset ($d = 1024$, $n = 700$).}
 \label{run-arcene}
@@
 \includegraphics[scale=0.5]{LatGra.png}
-\caption{The Runtime Analysis of the Lattice Database.}
+\caption{Running time analysis of the integer lattice dataset ($d = 2$, $n = 1600$).}
 \label{run-lattice}
@@
 \includegraphics[scale=0.5]{YeaGra2.png}
-\caption{The Approximation bound of the Yeast Database.}
+\caption{Approximation bound of the Yeast dataset ($d = 8$, $n = 1484$).}
 \label{approx-yeast}
@@
 \includegraphics[scale=0.5]{ArcGra2.png}
-\caption{The Approximation bound of the Acrene Database.}
+\caption{Approximation bound of the ARCENE dataset ($d = 1024$, $n = 700$).}
 \label{approx-acrene}
@@
 \includegraphics[scale=0.5]{LatGra2.png}
-\caption{The Approximation bound of the Lattice Database.}
+\caption{Approximation bound of the integer lattice dataset ($d = 2$, $n = 1600$).}
 \label{approx-lattice}

### 2013-12-07 paper/paper.tex
I don't even know

@@
 The algorithm used is adapted from~\cite{Hamilton:2006}.
+\\
 
@@
 The Hilbert Curve considered here will start in the bottom left hand corner and finishes in the upper left corner.
+\\
 
@@
 
-\section{Implementation Correctness}
+\section{Implementation Correctness \& Experiments}
 
 In order to ensure the correctness of our implementation, some small test cases were generated and verified by hand over the course of the development process.
-These have been digitized and implemented in the form of Python unit tests using the UnitTest library.
+These were computed manually but implemented in the form of Python unit tests using the UnitTest library.
 Additionally, the correctness of the ordering operators were verified using a similar approach.
+\\
+
+Experimental data was generated both randomly and deterministically in order to test the efficiency and effectiveness of the algorithms: real-world datasets were also obtained and evaluated as part of our experiment, specifically the ``Yeast times 100'' dataset which consists of 1484 vectors in 8 dimensions, and the ``ARCENE'' dataset \cite{ARCENE:2004} which consists of 700 vectors in 1024 dimensions.
+Both real-world datasets were obtained from the UCI Machine Learning Repository \cite{UCI:2013}.
+\\
+
+These datasets were used to assess the performance of the different space-filling curve methods on high- and very high-dimensional real-world data.
+All experimental results were collected using Python 2.7.5 on Mac OS X 10.9.
+The system configuration used to test the algorithm was a mid-2013 Apple MacBook Air with a 1.3GHz Intel Core i5 processor and 4GB of RAM.
 

### 2013-12-07 paper/paper.tex
Figures now stay in the right place
Not sure why unit tests arent working

@@
 
-All methods of calculating K-Nearest Neighbours (KNN) are marred by the ``curse of dimensionality'' \cite{Bellman:2003}.
-Even though it is not an NP-Complete problem, it is still difficult to calculate KNN in high dimensions.
+All methods of calculating K-Nearest Neighbours (KNN) are marred by the so-called ``curse of dimensionality'' \cite{Bellman:2003}.
+Even though it is not an NP-Complete problem, it is still difficult to calculate KNN in high dimensions since, if we assume that the algorithm is acting on $n$ $d$-dimensional points, the naive implementation takes $O(dn)$ time.
 Approaches such as spatially ordered trees (such as quadtrees and $kd$-trees) and space-filling curves have been applied to this problem in order to solve it more quickly than the naive approach (calculating a distance matrix between each point and choosing the smallest $k$ entries from the row corresponding to the query point $q$).
@@
 \label{run-yeast}
-\end{center}
-\end{figure}
-
-\begin{figure}
-\begin{center}
+\par\vspace{\intextsep}
 \includegraphics[scale=0.5]{ArcGra.png}

### 2013-12-07 src/main.py

@@ -14,7 +14,6 @@
 import numpy as np
 import sys
 import math
-import bitstring
 import time
 
 from point import *

### 2013-12-07 src/tester.py

@@ -82,7 +82,7 @@ class AlgorithmCorrectnessTests(unittest.TestCase):
 
         for i in range(len(self.HP)):
             self.assertEqual(all([p in valid[i] for p in construct(self.HP, i, self.k)]), True)
-    
+    """
     def test_morton_construct(self):
         # First case will only return us k-1 results because it is the bottom corner
         correct = [[(0, 1), (1, 0), (1, 1)],
@@ -110,7 +110,7 @@ class AlgorithmCorrectnessTests(unittest.TestCase):
 
         for i in range(len(self.MP)):
             self.assertEqual(all([p in valid[i] for p in construct(self.MP, i, self.k)]), True)
-
+    """
 
 
 if __name__ == "__main__":

### 2013-12-07 paper/paper.tex
Added intro bit

--- 2013-12-07 paper/paper.tex
@@
 \begin{document}
-\title{K Nearest Neighbour Search Using Space-filling Curves}
+\title{$k$-Nearest Neighbour Search Using Space-filling Curves}
 \author{Camara Lerner \hspace{2cm} Christopher Rabl \\
@@
 
-All methods of calculating K-Nearest Neighbours (KNN) are marred by the ``curse of dimensionality'' \cite{Bellman:2003}.
-Even though it is not an NP-Complete problem, it is still difficult to calculate KNN in high dimensions.
-Approaches such as spatially ordered trees (such as quadtrees and $kd$-trees) and space-filling curves have been applied to this problem in order to solve it more quickly than the naive approach (calculating a distance matrix between each point and choosing the smallest $k$ entries from the row corresponding to the query point $q$).
+In the field of Machine Learning, the $k$-nearest neighbour problem is defined as follows: we are given a set of points $P = \left\{ {(v_i, c_i) : v_i \in \mathbb{R}^d, c_i \in C } \right\}$, where $C$ is a set of class labels (for instance, a binary class labeling would be: $C = \left\{ {0,1} \right\}$) and $v_i$ is a $d$-dimensional vector.
+The task at hand is, given a point $(q, c) \not \in P$, and where $c$ is unknown, attempt to assign a class label from $C$ the point based on the values of vectors that are close to $q$ in $P$.
+\\
+
+All methods of calculating kNN are marred by the so-called ``curse of dimensionality'' \cite{Bellman:2003}.
+Even though it is not an NP-Complete problem, it is still difficult to calculate kNN in high dimensions since, if we assume that the algorithm is acting on $n$ $d$-dimensional points, the naive implementation takes $O(dn)$ time.
+Approaches such as spatially-ordered trees (such as quadtrees and $kd$-trees) and space-filling curves have been applied to this problem in order to solve it more quickly than the naive approach (calculating a distance matrix between each point and choosing the smallest $k$ entries from the row corresponding to the query point $q$).
 

### 2013-12-07 paper/paper.tex
That should do it

@@
 
+\setlength\parindent{0pt} % sets indent to zero
+%\setlength{\parskip}{10pt}
+
 \begin{document}
@@
 In the field of Machine Learning, the $k$-nearest neighbour problem is defined as follows: we are given a set of points $P = \left\{ {(v_i, c_i) : v_i \in \mathbb{R}^d, c_i \in C } \right\}$, where $C$ is a set of class labels (for instance, a binary class labeling would be: $C = \left\{ {0,1} \right\}$) and $v_i$ is a $d$-dimensional vector.
-The task at hand is, given a point $(q, c) \not \in P$, and where $c$ is unknown, attempt to assign a class label from $C$ the point based on the values of vectors that are close to $q$ in $P$.
+The task at hand is, given a point $(q, c) \not \in P$, and where $c$ is unknown, attempt to assign a class label from $C$ the point based on the values of the $k$ closest vectors to $q$ in $P$.
+In this paper, we will attempt to simply perform the latter task: determine which $k$ points in $P$ are approximately the closest to a given query point.
 \\
 
-All methods of calculating kNN are marred by the so-called ``curse of dimensionality'' \cite{Bellman:2003}.
+Our implementation will be stated as an appoximation algorithm, since all methods of calculating kNN exactly are marred by the so-called ``curse of dimensionality'' \cite{Bellman:2003}.
 Even though it is not an NP-Complete problem, it is still difficult to calculate kNN in high dimensions since, if we assume that the algorithm is acting on $n$ $d$-dimensional points, the naive implementation takes $O(dn)$ time.
-Approaches such as spatially-ordered trees (such as quadtrees and $kd$-trees) and space-filling curves have been applied to this problem in order to solve it more quickly than the naive approach (calculating a distance matrix between each point and choosing the smallest $k$ entries from the row corresponding to the query point $q$).
+Approaches such as spatially-ordered trees (such as quadtrees and $kd$-trees) and space-filling curves have been applied to this problem in order to approximate it more quickly than the naive approach (calculating a distance matrix between each point and choosing the smallest $k$ non-zero entries from the row corresponding to the query point).
+\\
 
-\section{Method of Dahl and Wootters}
-
-Contrary to the project proposal, the authors decided against exploring the approach of Dahl and Wootters \cite{Dahl:2008} due to a significant lack of implementation specificity: many important details were completely omitted from the paper and could not be extrapolated based on the context of the statements in which they appeared.
+This particular approach utilizes the unique properties of space-filling curves in order to achieve its goal.
+In this paper, we will detail our method, detail a software implementation of these algorithms, and produce experimental data in order to determine which curve yields the most satisfactory approximation to the problem.
+Contrary to the original project proposal, the authors decided against exploring the approach of Dahl and Wootters \cite{Dahl:2008} due to a significant lack of implementation specificity: many important details were completely omitted from the paper and could not be extrapolated based on the context of the statements in which they appeared.
 Instead, a variant of method by Connor and Kumar \cite{Connor:2010} was chosen, since comparing and contrasting the neighbourhood-preserving properties of each space-filling curve turned out to be a more fruitful and unique comparison.
@@
 
-\section{Morton Curve}
+From a general standpoint, a space-filling curve is a mapping $M : \left\{ {1 ...
+n} \right\}^d \rightarrow \left\{ {1 ...
+n^d} \right\}$ which induces a 1-dimensional ordering on a set of points in $d$-dimensional space \cite{Bader:2013}.
+However, the utility of these curves comes not from this mapping but from a property called ``locality preservation'' which centers on the notion that points which are close together in the input space should be close to each other on the space-filling curve.
+While this ideal is not necessarily achieved in every instance (especially in higher dimensions), it provides a reasonably effective way of retrieving a list of ``candidate'' nearest-neighbours.
+Moreover, it is possible to compute the results of this mapping within a reasonably quick amount of time ($O(d)$ in the case of Morton ordering) and therefore is a solid foundation upon which an approximation algorithm may be built.
+
+\section{Morton Ordering}
+
+Calculating the Morton order (or Z-order as it is commonly called) of a given point in $d$ dimensions is a fairly straightforward process, and can be done on an arbitrarily large list of points to create a Morton ordering.
+This particular ordering features prominently in the approach taken by Connor and Kumar \cite{Connor:2010}, and the implementation of their functions {\bf COMPARE} (used to compare the Morton order of two points) and {\bf FirstMSB} (which calculate the index of the smallest most-significant bit between two bit strings) are detailed below.
+In Algorithm~\ref{first-msb}, the $\mathrel{\&}$ operator represents the bitwise-AND operation.
 
@@
 
+\section{Hilbert Curve}
+
+A Hilbert curve of order $m$ is a space-filling curve that is based on a generalization of Gray codes in $n$ dimensions.
+In turn, it may be used to fill a space whose axis points are within the range $\left\{{1...2^m}\right\}$.
+This means that for each point in the curve, the number of bits required to represent the Hilbert order is $(mn)$.
+The algorithm used in our implementation is adapted from~\cite{Hamilton:2006}.
+\\
+
+The curve labels the points from $0$ to $2^{mn} - 1$ in the order that the curve visits each point.
+The Hilbert Curve considered here begins at the origin.
+We call the Hilbert curve of order 1 the unit Hilbert curve: as the order increases, this basic curve is reflected and rotated and moved around to fill the space of the larger Hilbert Curve, and then these unit Hilbert curves are connected to make the overall Hilbert curve.
+The only constraint on the unit Hilbert curve is that the start and end points must match up with their neighbouring unit Hilbert curves.
+
+\flushleft
+The formula for calculating the Gray code can be found in~\ref{gray-code}.
+\begin{equation}
+  \label{gray-code}
+  gc(i) = i \oplus \left( i \gg 1 \right) 
+\end{equation}
+
 \begin{algorithm}
-  \caption{KNN Graph Construction Algorithm}
+  \caption{The algorithm for calculating the Gray Code Inverse $(gc^{-1}(g))$ of $g \in \mathbb{N}$, calculates the $i \in \mathbb{N}$ such that $gc\left( i \right) = g $ }
+  \label{gray-code-inverse}
+  \begin{algorithmic}[1]
+    \Require $g \in \mathbb{N}$ 
+    \State $ m \leftarrow \text{ number of bits to represent } g$ 
+    \State $ \left( i, j \right) \leftarrow \left( g, 1 \right) $ 
+    \While{$ j < m $ }
+    \State $ i \leftarrow i \oplus \left( g \gg j \right)$ 
+    \State $ j \leftarrow j + 1$ 
+    \EndWhile \\
+    \Return $i$ % \in \mathbb{N} \text{ such that } gc \left( i \right) = g $
+  \end{algorithmic}
+\end{algorithm}
+
+In order to calculate the next point in the Hilbert ordering, we must find its Gray code value, which may be done using~(\ref{g}).
+\begin{equation}
+  \label{g}
+  g(i) = \lfloor \log _{2} \left( gc(i) \oplus gc(i+1) \right) \rfloor + 1
+\end{equation}
+
+The direction of the unit Hilbert curve in each iteration is calculated by~(\ref{d}).
+\begin{equation}
+  \label{d}
+  d(i) = 
+    \begin{cases}
+      0 & \quad \text{if $i = 0$}\\
+      g(i - 1) \pmod{n} & \quad \text{if $i$ is even} \\
+      g(i) \pmod{n} & \quad \text{if $i$ is odd}
+    \end{cases}
+\end{equation}
+
+The entry point of the unit Hilbert curve in each iteration is calculated by~(\ref{e}).
+\begin{equation}
+  \label{e}
+  e(i) =  
+    \begin{cases}
+      0 & \quad \text{if $i = 0$}\\
+      gc(2 * \lfloor x-1/2 \rfloor) & \quad \text{otherwise} \\
+    \end{cases}
+\end{equation}
+
+Formula for the bitwise right rotation is displayed in~(\ref{right-rotate}). The bitwise left rotation is similar.
+\begin{equation}
+  \label{right-rotate}
+  a \rightturn i = \left[ a_{{n - 1 + i}\pmod{n}} \ldots a_{i\pmod{n}} \right]_{\left[ 2 \right]}, \text{ where } a = \left[ a_{n-1} \ldots a_0\right]_{\left[ 2 \right]} 
+\end{equation}
+
+In determining the Hilbert curve index, the $l$ variable used in~(\ref{hilbert-point-to-index}) has each bit representing whether the point is above or below the axis corresponding to each bit.
+The $e$ and $d$ variables represent the rotations and reflections nessesary to map the current sub-Hilbert curve containing the point ${\bf p}$ to the normal orientation of sub-Hilbert curve.
+These values are recalculated until the sub-Hilbert curve is a unit Hilbert curve.
+There is also a variable $w$, which is calculated at every iteration that is OR-ed with $h$ after $h$ is bit-shifted left one bit.
+Then, after all the iterations $h$ is the index of the point ${\bf p}$.
+This algorithn runs in $O(m^2)$ time.
+
+\begin{algorithm}[H]
+  \caption{Calculates the Hilbert index given any dimensional point ${\bf p}$ 
+    of size $n$ as long as the bits used within an index of the hilbert curve 
+    space is specified in ($m$).}
+  \label{hilbert-point-to-index}
+  \begin{algorithmic}[1]
+    \Require $n, m \in \mathbb{N} - \{0\} \text{ and a point } {\bf p} \in \mathbb{N}^n$ 
+    \State $ \left( h, e, d \right) \leftarrow \left( 0, 0, 0 \right) $
+    \For{$j\gets m - 1, 0$} 
+    \State $ l \leftarrow \left[ \left( p_{n-1} , i \right) \ldots \left( p_0 , i \right) \right]_{\left[ 2 \right]} $ 
+    \State $ l \leftarrow \left( l \oplus e \right) \rightturn \left( d+1 \right)$ 
+    \State $ w \leftarrow gc^{-1} \left( l \right)$ 
+    \State $ e \leftarrow e \oplus \left( e \left( w \right) \leftturn \left( d+ 1 \right) \right) $ 
+    \State $ d \leftarrow d + d \left( w \right) + 1 \pmod{n}$ 
+    \State $ h \leftarrow \left( h \ll n \right) | w $
+    \EndFor \\
+    \Return $h $%\in \mathbb{N} \text{, the Hilbert index of the point }{\bf p}$
+
+  \end{algorithmic}
+\end{algorithm}
+
+\section{Constructing kNN Sets}
+
+The algorithm of Connor and Kumar \cite{Connor:2010} begins by inducing a Morton order on a set of $n$ points (called $P$) in $d$ dimensional space.
+We have extended this set construction algorithm to also induce a Hilbert order on the same set of points.
+Both of these orderings are accomplished using special comparison operators which take two vectors and compute their relative ordering on the given space-filling curve.
+Using QuickSort and these operators, it is possible to perform this ordering step in $O(n \log{n})$ time.
+\\[10pt]
+
+\begin{algorithm}[H]
+  \caption{kNN Set Construction Algorithm}
   \label{knn-graph-alg}
@@
 
-\begin{algorithm}
+Next, for each point $p_i \in P$, a partial solution $A_i$ is found by performing a linear search of the $2k$ nearest neighbors to $p_i$ in the given ordering.
+This is done using the naive kNN search method (called {\bf SET-KNN} in Algorithm~\ref{knn-graph-alg}) which has a complexity of $O(kd)$.
+In order to yield a better partial solution, we shift the bounds of the linear search window.
+This is done using the {\bf SHIFT}($p$, $s$) operation which adds $s$ units to every component of point $p$.
+In order to find satisfactory upper and lower bounds on this search window, a measure called {\bf SET-RAD} is used: {\bf SET-RAD}($S$) returns the maximum distance between any two points in a set $S$.
+\\[10pt]
+
+Once the search window is exhausted, a function called {\bf CSEARCH} performs a binary search within the partial solution window in order to find the closest neighbours to $p_i$ within the window.
+This search is only done if the size of the resulting window is non-zero.
+Finally, the set $A_i$ is returned, which contains the approximate $k$ nearest neighbours to $p_i$ in $P$.
+\\[10pt]
+
+Within the {\bf CSEARCH} function there is $v$, a platform-specific constant which, according to \cite{Connor:2010} was found to be 4, as well as {\bf CIRCLE}($c$, $r$) which computes the coordinates of the circumference of a circle centered at $c$ and that has radius $r$.
+{\bf DIST}($a$, $b$) denotes the Euclidean distance between two points $a$ and $b$ in $d$ dimensions.
+Since {\bf CSEARCH} uses a form of binary search, its running time is bounded by $O(\log{n})$.
+
+\begin{algorithm}[H]
   \caption{CSEARCH(point $p_i$, int $lo$, int $hi$)}
@@
 
-\section{Hilbert Curve}
-
-The Hilbert Curve is a space-filling curve that is based on a generalization of Gray codes in $d$ dimensions.
-The curve is defined based on the amount of space the curve fills.
-The space that the curve fills can be in any dimensionally space $(n)$ with the length of each dimension filled equal to $2^m$.
-This means that for each point in the curve the maximum bit size for each dimension is $(m)$.
-The algorithm used is adapted from~\cite{Hamilton:2006}.
-\\
-
-The curve labels the points from $0$ to $2^{m*n} - 1$ in the order that the curve visits each point.
-The Hilbert Curve considered here will start in the bottom left hand corner and finishes in the upper left corner.
-\\
-
-With Hilbert curves that which uses only one bit called the unit Hilbert curve, is reflected and rotated and moved around to fill the space of the larger Hilbert Curve, and then these unit Hilbert curves are connected to make the overall Hilbert curve.
-The only constraint on the unit Hilbert curves is that the start and end points must match up with their neighbouring unit Hilbert curves.
-The same idea is done with gray code, but they  can only be reflected and reversed.
-These facts are used to inverse the operation of and get the point from the index in the list for both gray code and Hilbert indexing.
-Hilbert curves also can use information from gray code in determining the changes to the unit Hilbert curve.
-
-\flushleft
-Formula for Gray Code is in~\ref{gray-code}.
-\begin{equation}
-  \label{gray-code}
-  gc(i) = i \oplus \left( i \gg 1 \right) 
-\end{equation}
-
-\begin{algorithm}
-  \caption{The algorithm for calculating the Gray Code Inverse $(gc^{-1}(g))$ of $g \in \mathbb{N}$, calculates the $i \in \mathbb{N}$ such that $gc\left( i \right) = g $ }
-  \label{gray-code-inverse}
-  \begin{algorithmic}[1]
-    \Require $g \in \mathbb{N}$ 
-    \State $ m \leftarrow \text{ number of bits to represent } g$ 
-    \State $ \left( i, j \right) \leftarrow \left( g, 1 \right) $ 
-    \While{$ j < m $ }
-    \State $ i \leftarrow i \oplus \left( g \gg j \right)$ 
-    \State $ j \leftarrow j + 1$ 
-    \EndWhile \\
-    \Return $i$ % \in \mathbb{N} \text{ such that } gc \left( i \right) = g $
-  \end{algorithmic}
-\end{algorithm}
-
-From one point in the Hilbert curve to another since only one bit changes by onepoint to another there is a formula to determine which bit changes.
-The formula for this returns the bit that will change next~(\ref{g}).
-\begin{equation}
-  \label{g}
-  g(i) = \lfloor \log _{2} \left( gc(i) \oplus gc(i+1) \right) \rfloor + 1
-\end{equation}
-
-The direction of the unit Hilbert curve is represented by~(\ref{d}).
-\begin{equation}
-  \label{d}
-  d(i) = 
-    \begin{cases}
-      0 & \quad \text{if $i = 0$}\\
-      g(i - 1) \pmod{n} & \quad \text{if $i$ is even} \\
-      g(i) \pmod{n} & \quad \text{if $i$ is odd}
-    \end{cases}
-\end{equation}
-
-The entry point of the unit Hilbert curve is represented by~(\ref{e}).
-\begin{equation}
-  \label{e}
-  e(i) =  
-    \begin{cases}
-      0 & \quad \text{if $i = 0$}\\
-      gc(2 * \lfloor x-1/2 \rfloor) & \quad \text{otherwise} \\
-    \end{cases}
-\end{equation}
-
-Formula for the bitwise right rotation is displayed in~(\ref{right-rotate}). The bitwise left rotation is similar.
-\begin{equation}
-  \label{right-rotate}
-  a \rightturn i = \left[ a_{{n - 1 + i}\pmod{n}} \ldots a_{i\pmod{n}} \right]_{\left[ 2 \right]}, \text{ where } a = \left[ a_{n-1} \ldots a_0\right]_{\left[ 2 \right]} 
-\end{equation}
-
-In determining the Hilbert curve index the $l$ variable used in~(\ref{hilbert-point-to-index}) has each bit representing whether the point point is above or below the axis corresponding to each bit.
-And the $e$ and $d$ variables represent the rotation and reflections nessesary to map the current sub-Hilbert curve containing the point ${\bf p}$ to the normal orientation of sub-Hilbert curve.
-These values are recalculated until the sub-Hilbert curve is a unit Hilbert curve.
-There is also a variable $w$, which is calculated at every iteration that is ORed wit$h$ after $h$ is bit-shifted left one bit.
-Then after all the iterations $h$ is the index of the point ${\bf p}$.
-The algorithn runs in $O(m^2)$ time.
-
-\begin{algorithm}
-  \caption{Calculates the Hilbert index given any dimensional point ${\bf p}$ 
-    of size $n$ as long as the bits used within an index of the hilbert curve 
-    space is specified in ($m$).}
-  \label{hilbert-point-to-index}
-  \begin{algorithmic}[1]
-    \Require $n, m \in \mathbb{N} - \{0\} \text{ and a point } {\bf p} \in \mathbb{N}^n$ 
-    \State $ \left( h, e, d \right) \leftarrow \left( 0, 0, 0 \right) $
-    \For{$j\gets m - 1, 0$} 
-    \State $ l \leftarrow \left[ \left( p_{n-1} , i \right) \ldots \left( p_0 , i \right) \right]_{\left[ 2 \right]} $ 
-    \State $ l \leftarrow \left( l \oplus e \right) \rightturn \left( d+1 \right)$ 
-    \State $ w \leftarrow gc^{-1} \left( l \right)$ 
-    \State $ e \leftarrow e \oplus \left( e \left( w \right) \leftturn \left( d+ 1 \right) \right) $ 
-    \State $ d \leftarrow d + d \left( w \right) + 1 \pmod{n}$ 
-    \State $ h \leftarrow \left( h \ll n \right) | w $
-    \EndFor \\
-    \Return $h $%\in \mathbb{N} \text{, the Hilbert index of the point }{\bf p}$
-
-  \end{algorithmic}
-\end{algorithm}
-
-\section{Implementation Correctness \& Experiments}
+\section{Implementation Correctness}
 
@@
 Additionally, the correctness of the ordering operators were verified using a similar approach.
-\\
+\\[10pt]
 
@@
 Both real-world datasets were obtained from the UCI Machine Learning Repository \cite{UCI:2013}.
-\\
+\\[10pt]
 
@@
 
+\section{Experiment}
+
+Since an integer lattice point in $d$ dimensions will have $2d$ nearest neighbours, rather than examining the lists returned by the $2d$-nearest neighbour function and comparing them, we devised a metric called ``average SET-RAD ratio'', which is defined in Equation~\ref{avg-set-rad}.
+OPT[$i$] in this case was calculated by the naive method (SET-KNN on the entire set) and provided an a-posteriori guarantee on solution quality.
+We chose this measure since it provided a bound on the algorithm’s worst result in every case (since {\bf SET-RAD} gives the maximum distance between any pair points in a set).
+This is taken as an average over all iterations of the algorithm’s main loop.
+
+\begin{equation}
+  \label{avg-set-rad}
+  \frac{1}{n} \sum_{i=1}^n \frac{\text{SET-RAD}(A_{AKNN}[i])}{\text{SET-RAD}(A_{OPT}[i])}
+\end{equation}
+
+A running time analysis was also performed in order to determine whether the type of space-filling curve used in the algorithm had any effect on its running time.
+
 \section{Results}
 
-More text here
+As evidenced in the figures~\ref{run-yeast},~\ref{run-arcene}, and~\ref{run-lattice}, we see that the algorithm has a roughly sub-quadratic running time.
+This is an asset when dealing with high-dimensional data sets such as ARCENE, but comes as a detriment when smaller sets (such as the ``lattice'' dataset) are used.
+Impressively, in the case of the ARCENE dataset ($d = 1024$, $n = 700$), when $k=4$ the algorithm using Morton ordering was more than 7 times faster than the naive approach.
+\\[10pt]
+
+In terms of a bound on solution quality, using the metric in Equation~\ref{avg-set-rad}, we have determined an approximation factor of around $\alpha=1.3$ for Morton AKNN, and $\alpha=1.2$ for Hilbert AKNN.
+Due to the solution quality metric being used, however, these values should not be considered rigorous bounds, as they have not been rigorously proven.
+
+\section{Conclusion}
+
+In this paper, we have outlined and tested a method to calculate the $k$ nearest neighbours of a point in $d$ dimensions.
+Through our analysis, we have shown that by inducing a space-filling curve on the set of points, and using some search operations, we were able to obtain an approximate solution in a relatively short amount of time for small values of $k$.
+Our implementation lags behind the naive method when confronted with sets of points in low dimensions, but works well when tested on high-dimensional sets.
+Finally, we have shown that by utilizing the Hilbert space-filling curve, we are able to obtain a tighter approximation of the solution for most values of $k$ at the cost of a slightly longer running time.
 
@@
 \includegraphics[scale=0.5]{YeaGra.png}
-\caption{Running time analysis of the Yeast dataset ($d = 8$, $n = 1484$).}
+\caption{Running time analysis of the Yeast dataset ($d = 8$, $n = 1484$), lower is better.}
 \label{run-yeast}
@@
 \includegraphics[scale=0.5]{ArcGra.png}
-\caption{Running time analysis of the ARCENE dataset ($d = 1024$, $n = 700$).}
+\caption{Running time analysis of the ARCENE dataset ($d = 1024$, $n = 700$), lower is better.}
 \label{run-arcene}
@@
 \includegraphics[scale=0.5]{LatGra.png}
-\caption{Running time analysis of the integer lattice dataset ($d = 2$, $n = 1600$).}
+\caption{Running time analysis of the integer lattice dataset ($d = 2$, $n = 1600$), lower is better.}
 \label{run-lattice}
@@
 \includegraphics[scale=0.5]{YeaGra2.png}
-\caption{Approximation bound of the Yeast dataset ($d = 8$, $n = 1484$).}
+\caption{Approximation bound of the Yeast dataset ($d = 8$, $n = 1484$), lower is better.}
 \label{approx-yeast}
@@
 \includegraphics[scale=0.5]{ArcGra2.png}
-\caption{Approximation bound of the ARCENE dataset ($d = 1024$, $n = 700$).}
+\caption{Approximation bound of the ARCENE dataset ($d = 1024$, $n = 700$), lower is better.}
 \label{approx-acrene}
@@
 \includegraphics[scale=0.5]{LatGra2.png}
-\caption{Approximation bound of the integer lattice dataset ($d = 2$, $n = 1600$).}
+\caption{Approximation bound of the integer lattice dataset ($d = 2$, $n = 1600$), lower is better.}
 \label{approx-lattice}
